---
title: "UBCF Movie Recommender model"
---

```{r}
library("recommenderlab")

data("MovieLense")
```

Only use users with more than 100 ratings.
```{r}
MovieLense100 <- MovieLense[rowCounts(MovieLense) >100,]

MovieLense100
```

Train UBCF recommender with small training set of 50 users.
```{r}
train <- MovieLense100[1:50]

rec <- Recommender(train, method = "UBCF")
rec
```

Generate top-N recommendations for new users (users 101 & 102) as per the documentation.
```{r}
str(MovieLense100[101])
head(as(MovieLense100[101], "data.frame"),9)
pre <- predict(rec, MovieLense100[101:102], n = 5)
pre
str(pre)
as(pre, "list")
```

Now lets try to get some recommendations for our own film ratings. 

First, create a vector of movie names.
```{r}
films <- MovieLense100[100]@data@Dimnames[[2]]
```
Everything is better in alphabetical order.
```{r}
films <-films[order(films)]
head(films)
```
Then create a sparse matrix.
```{r}
m<-matrix(0,ncol = length(films))
colnames(m) <- films
m[,1:5]

m
```

Here is a vector of some films:
```{r}
mymovies<-c("Toy Story (1995)","Four Rooms (1995)","Get Shorty (1995)","Copycat (1995)",
            "Twelve Monkeys (1995)","Babe (1995)","Dead Man Walking (1995)","Seven (Se7en) (1995)",
            "Usual Suspects, The (1995)","Absolute Power (1997)","Abyss, The (1989)","Young Guns II (1990)")
```
Here are the ratings:
```{r}
myratings<-c(5,3,4,3,5,4,1,2,2,4,5,3)
```

Put the ratings in alphabetical order of mymovies.
```{r}
myratings <-myratings[order(mymovies)]
```
Put mymovies in alphabetical order.
```{r}
mymovies <-mymovies[order(mymovies)]
```
Clever substitution of myratings into the matrix.
```{r}
m[,which(films %in% mymovies)] <- myratings
```
Create a realRatingMatrix.
```{r}
m <- as(m, "sparseMatrix")
m
r <- as(m, "realRatingMatrix")
pre <- predict(rec, r, n = 20)
as(pre,"list")
```

Now try this to randomly select 100 movies and give random ratings between 1 and 5.
```{r}
N=100
mymovies <- sample(films,N)
myratings <- sample(1:5,N,replace=TRUE)
```

Put the ratings in alphabetical order of mymovies.
```{r}
myratings <-myratings[order(mymovies)]
```
Put mymovies in alphabetical order.
```{r}
mymovies <-mymovies[order(mymovies)]
```
Cleverly substitute myratings into the matrix.
```{r}
m[,which(films %in% mymovies)] <- myratings
```
Create a realRatingMatrix.
```{r}
m <- as(m, "sparseMatrix")
m
r <- as(m, "realRatingMatrix")
pre <- predict(rec, r, n = 20)
as(pre,"list")
```
