---
title: "Yeast Clusters with Gene Ontology"
author: "MSoria"
date: "8/24/2020"
output: html_document
---

</br>

***

##### The following code associates annotations from Gene Ontology to different clusters from the yeast genome using data from DEE2  

***

</br>


### Prerequisite libraries
```{r Libraries, echo=TRUE}
suppressPackageStartupMessages({c(library(R.utils), library(dplyr), library(tidyr), library(data.table), library(RColorBrewer), library(gplots), library(Polychrome), library(tidyverse), library(edgeR), library(GO.db), library(gage), library(gProfileR))})

```
</br>

###  1. Download and unzip the Yeast data 
```{r Download Data, echo=TRUE}
# put in an if statement to check if object exists

if (!file.exists("Data/scerevisiae_se.tsv.bz2")){
    download.file("http://dee2.io/mx/scerevisiae_se.tsv.bz2", 
              destfile = "Data/scerevisiae_se.tsv.bz2")}

# fread error msg: This file is very unusual: it ends abruptly without a final newline, and also its size is a multiple of 4096 bytes. Please properly end the last row with a newline using for example 'echo >> file' to avoid this  time to copy. 

# needed to go back to bunzip2
if (!file.exists("Data/scerevisiae_se.tsv")){
    bunzip2("Data/scerevisiae_se.tsv.bz2", destname="Data/scerevisiae_se.tsv")}

scerevisiae_data <- fread("Data/scerevisiae_se.tsv", sep = '\t', header = FALSE)

head(scerevisiae_data)
```
</br>

### 2. Download, read, and unzip the QC data
```{r QC Data, echo=TRUE}
if (!file.exists("Data/scerevisiae_qc.tsv.bz2")){
    download.file("http://dee2.io/mx/scerevisiae_qc.tsv.bz2", 
              destfile = "Data/scerevisiae_qc.tsv.bz2")}

quality_data <- fread("Data/scerevisiae_qc.tsv.bz2", sep = '\t', col.names = c("Database", "Description", "Details"))

head(quality_data)
```
</br>

### 3. Filter and quantify databases according to QC_SUMMARY = PASS.  
```{r QC Data Summary, echo=TRUE}

# count the total number of QC_SUMMARY entries for checking
qc_summary <- quality_data[quality_data$Description == 'QC_SUMMARY',]
total_summary <- nrow(qc_summary)

# count the total number of databases marked as PASS
qc_pass <- filter(quality_data, grepl("PASS", Details))
total_pass <- nrow(qc_pass)

```
</br>

### 4. Select all the databses from scerevisiae_se.tsv with QC_SUMMARY == 'PASS' 
```{r Filter yeast bulk data, echo=TRUE}

# filter all the sample names with QC_SUMMARY == 'PASS' from step 3
databse_pass <- filter(quality_data, grepl("PASS", Details)) %>% select(Database)

# convert the rows into string and store on a list to use for filtering Yeast data
databse_pass_list <- as.list(as.character(databse_pass$Database))

# Filter Yeast data using generated list
scerevisiae_pass <- scerevisiae_data[scerevisiae_data$V1 %in% databse_pass_list,] 

# Convert format from long to wide
scerevisiae_pass_wide <- scerevisiae_pass %>% pivot_wider(names_from = "V1", values_from = "V3")

# Convert tibble to data frame and assign column 1 as rowname
scerevisiae_pass_wide <- as.data.frame(scerevisiae_pass_wide)
rownames(scerevisiae_pass_wide) <- scerevisiae_pass_wide[,1]
scerevisiae_pass_wide <- scerevisiae_pass_wide[,-1]

scerevisiae_pass_wide[1:10,1:6]

```
</br>

### 5. Aggregate the SRR run data to SRX experiment 
```{r SRR to SRX Aggregation, echo=TRUE}

# Download and read Metadata summary

if (!file.exists("Data/scerevisiae_metadata.tsv.cut")){
    download.file("http://dee2.io/metadata/scerevisiae_metadata.tsv.cut", 
              destfile = "Data/scerevisiae_metadata.tsv.cut")}
    
yeast_metadata <- read.csv("Data/scerevisiae_metadata.tsv.cut", sep = '\t')

# Filter metadata to include only "passed" samples
yeast_metadata_pass <- yeast_metadata[yeast_metadata$SRR_accession %in% databse_pass_list,]

# Assign column 1 as rownames 
rownames(yeast_metadata_pass) <- yeast_metadata_pass[,1]
yeast_metadata_pass <- yeast_metadata_pass[,-1]

# Sequencing Run aggregate function
srx_agg <- function(x,counts="GeneCounts") {
    IDX=which(names(x) %in% "GeneCounts")
    mds<-x$MetadataSummary
    n=nrow(x[[IDX]])
    SRX_dat <- vapply(X=unique(mds$SRX_accession) ,function(srx) {
        srrs<-rownames(mds)[which(mds$SRX_accession %in% srx)]
        if (length(srrs)>1) {
            rowSums(x[[IDX]][,srrs])
        } else {
            x[[IDX]][,srrs]
        }
    } , numeric(n))
    rownames(SRX_dat) <- rownames(x[[IDX]])
    colnames(SRX_dat) <- unique(mds$SRX_accession)
    SRX_dat
}

# Put the filtered yeast GeneCount data and the filtered metadata dataframes into a list
yeast_count_metadata <- list(GeneCounts = scerevisiae_pass_wide, MetadataSummary = yeast_metadata_pass)

# Apply both dataframes to the function
agg <- srx_agg(yeast_count_metadata)

agg[1:10,1:6]

# Check for samples with zero total counts
colSums(agg) == 0

# Histogram showing samples (columns) counts
hist(colSums(agg), breaks = 30)

# Filtering samples with more than 1 million reads
agg_f_col <- agg[, which(colSums(agg) >= 1e6)]

# Filtering genes with more than zero counts 
agg_f_row <- agg[which(rowSums(agg) > 0), ]

# No difference between the filtered and unfiltered columns  
dim(agg_f_col)
dim(agg)

# Rows ommited for genes with more than zero counts
dim(agg_f_row)


```
</br>

### 6. Heirarchical Clustering with PASS only databases 

```{r eval=FALSE, include=FALSE}
# NORMALISATION FOR LIBRARY SIZE BIAS (cpm)
# Filtering to remove lowly expressed genes
myCPM <- cpm(agg)
thresh <- myCPM > 0.6

# Keep genes that have at least 2 TRUEs in each row of thresh
keep <- rowSums(thresh) >= 2

# Subset the rows of data to keep the more highly expressed genes
counts_keep <- agg[keep,]
summary(keep)
dim(counts.keep)

# A CPM of 0.6 is used as it corresponds to a count of 10-15 for the library sizes in this data set. If the count is any smaller, it is considered to be very low, indicating that the associated gene is not expressed in that sample. As a general rule, a good threshold can be chosen by identifying the CPM that corresponds to a count of 10, which in this case is about 0.6. You should filter with CPMs rather than filtering on the counts directly, as the latter does not account for differences in library sizes between samples.

# Note: When in doubt, a threshold of 1 CPM in at least minimum group sample size is a good rule of thumb.

# Check whether our threshold of 0.6 does indeed correspond to a count of about 10-15
# Add a vertical line at 0.6 CPM, and a horizontal line at Counts = 10
plot(myCPM[,1], agg[,1], ylim=c(0,50), xlim=c(0,3), abline(v=0.6, h=10, col="blue"), ylab = "Counts", xlab = "CPM", main = "Counts-per-million (CPM) vs Counts")

# Convert counts to DGEList object (from edgeR library)
dgeObj <- DGEList(counts_keep)

#NORMALISATION FOR COMPOSITION BIAS (TMM normalisation)
# Apply normalisation to DGEList object
dgeObj <- calcNormFactors(dgeObj)
# dgeObj$samples

# Get log2 counts per million
logcounts <- cpm(dgeObj,log=TRUE)


# dev.off()
```
</br>


### 8. Heirarchical Clustering with PASS only databases
```{r echo=TRUE}

# Applying Hierarchical Clustering after normalization
cl <-as.dist(1-cor(t(logcounts), method="spearman"))
hr <- hclust(cl , method="complete")

cl_mat <- as.matrix(cl)

# optimizing the cluster size
mycl <- cutree(hr, h=max(hr$height/1.611))
head(mycl)

if (!exists("P100")){
  P100 <- createPalette(100,  c("#ff0000", "#00ff00", "#0000ff"))}
mycl_length <- length(unique(mycl))

clusterCols <- P100[1:mycl_length]
myClusterSideBar <- clusterCols[mycl]
colfunc <- colorRampPalette(c("blue", "white", "red"))
if (!file.exists("Data/scerevisiae_se.tsv")){
  write.table(mycl,file="Data/scerevisiae_pass.txt",quote=F,sep="\t")}

# create a 5 x 5 inch png image
# png("Data/yeast_heatmaps_PASS_normalization2.png", width = 5*300, height = 5*300, res = 300, pointsize = 8)          

# create the heatmap
heatmap.2(logcounts, main="S.cerevisiae PASS Samples",  Rowv=as.dendrogram(hr),
          dendrogram="both", scale="column", col = colfunc(25), trace="none", 
          RowSideColors= myClusterSideBar, margins = c(5,5))

# dev.off()
```
</br>



### 9. Gene Ontology Gene Matrix 
```{r}
# Get the GO Annotation File (GAF) format 2.1 of yeast
if (!file.exists("Data/scerevisiae_sgd.gaf.gz")){
    download.file("http://current.geneontology.org/annotations/sgd.gaf.gz", 
              destfile = "Data/scerevisiae_sgd.gaf.gz")}
if (!file.exists("Data/scerevisiae_sgd.gaf")){
    gunzip("Data/scerevisiae_sgd.gaf.gz", destname="Data/scerevisiae_sgd.gaf")}

scerevisiae_GO <- fread("Data/scerevisiae_sgd.gaf", sep = '\t', header = FALSE)
head(scerevisiae_GO)

# Subset unique GeneIDs (V11) and GO IDs (V5) in a dataframe. Select the first gene synonym from column V11 and delete the rest. Use this as the GeneID column. Add another column with 1 in each row indicating the relationship of the Gene ID to the GO ID

scerevisiae_GO_matrix <- data.frame(unique(scerevisiae_GO[,c(11,5)]))
scerevisiae_GO_matrix$GeneID <- gsub("\\|.*", " ", scerevisiae_GO_matrix$V11)
scerevisiae_GO_matrix$V3 <- as.numeric(paste(1, scerevisiae_GO_matrix$V3))
scerevisiae_GO_matrix <- scerevisiae_GO_matrix[,-1]

# Convert the dataframe to wide format with Gene ID as column 1 and GO ID as column names 
scerevisiae_GO_matrix_wide <- scerevisiae_GO_matrix %>% pivot_wider(names_from = "V5", values_from = "V3")
scerevisiae_GO_matrix_wide <- as.data.frame(scerevisiae_GO_matrix_wide)
colnames(scerevisiae_GO_matrix_wide)[1] <- "GeneID"


# convert NA to Zero
scerevisiae_GO_matrix_wide[is.na(scerevisiae_GO_matrix_wide)] <- 0
head(scerevisiae_GO_matrix_wide)

sample_heatmap<-heatmap(as.matrix(scerevisiae_GO_matrix_wide[1:100,2:100]))

```
</br>

### 10. Sample Cluster analysis using Cluster 1
```{r}
# Prepare Cluster data frame
clusters <- as.data.frame(mycl)
colnames(clusters) <- "ClusterNumber"

# make a new column, GeneID, from rownames
clusters$GeneID <- rownames(clusters)

# remove rownames
rownames(clusters) <- c()


# Select all genes for cluster 
cluster1 <- clusters[clusters$ClusterNumber == 1,] %>% select(GeneID) 
cluster1_list <- as.list(cluster1$GeneID)

cluster2 <- clusters[clusters$ClusterNumber == 2,] %>% select(GeneID) 
cluster2_list <- as.list(cluster2$GeneID)


# look for cluster gene match in GO dictionary and subset nonzero columns 
cluster1_GO <- scerevisiae_GO_matrix_wide[scerevisiae_GO_matrix_wide$GeneID %in% cluster1_list,] 
cluster1_GO_nonzero <- cluster1_GO[, !apply(cluster1_GO == 0, 2, all)]
cluster1_GO_nonzero


# select GO_IDs based on threshold 
thresh <- nrow(cluster1_GO)*.60
cluster1_GO_thresh <- cluster1_GO_nonzero[, colSums(cluster1_GO_nonzero != 0) > thresh]

cluster1_GO_thresh

```

</br>

### 9. Gene Enrichment Analysis using goProfileR
```{r echo=TRUE}

# make a list for genes of interest
cluster_1 <- cl_mat[rownames(cl_mat) %in% cluster1_list,]
genesOfInterest <- rownames(cluster_1)

# Gene enrichment query
goResults_1 <- gprofiler(query = genesOfInterest, 
                     organism = 'scerevisiae',
                     hier_filtering = 'none')

# goResults <- gost(query = genesOfInterest, 
#                      organism = 'scerevisiae')
# goResults$result$parents


## Subset and reorder gProfiler results to only include columns of interest
gprofiler_results_oe_reordered <- goResults_1[, c("term.id", "domain", "term.name", "p.value", "overlap.size", "term.size", "intersection")]

## Order the results by p-adjusted value
gprofiler_results_oe_reordered <- gprofiler_results_oe_reordered[order(gprofiler_results_oe_reordered$p.value), ]

## Extract only the 'GO' terms from the results
gprofiler_results_oe_GOs <- gprofiler_results_oe_reordered[grep('GO:', gprofiler_results_oe_reordered$term.id), ]

## Write the enriched GO results to file
write.csv(gprofiler_results_oe_GOs, 
            "Data/gprofiler_cluster1.csv")

#-----------------------from gage Package--------------------------
# perform Gene Set Enrichment Analysis (GSEA) using gage package
library(org.Sc.sgd.db)
org.Sc.sgd()


columns(org.Sc.sgd.db)
keys(org.Sc.sgd.db)
keytypes(org.Sc.sgd.db)

# The input to gage must be Entrez identifiers so we first need to map our gene symbols to Entrez
ann <- mapIds(org.Sc.sgd.db, keys=rownames(cluster_1), column="ENTREZID", keytype="ENSEMBL")
cluster1_non_na <- names(!is.na(ann))
fc <- logcounts[rownames(logcounts) %in% cluster1_non_na,]

table(is.na(fc))

# Create a mapping between gene identifiers and GO terms.
allGo <- as.list(GOTERM)
bp <- Filter(function(go) go@Ontology == 'BP', allGo)
goMapping <- mget(names(bp), org.Sc.sgdGO2ALLORFS, ifnotfound=NA)

# gene set enrichment analysis
goFc <- gage(fc, goMapping)
goGreater <- as.data.frame(goFc$greater)
goGreater <- goGreater[goGreater$q.val < 0.01 & !is.na(goGreater$q.val),]
goTermNames <- lapply(mget(rownames(goGreater), GOTERM), function(go) go@Term)
goGreater$Term <- goTermNames[rownames(goGreater)]
goGreater

GSEA_heatmap <- heatmap(as.matrix(goGreater[,6:4840]))

goGreater[,1:6]
a$DRX021534
grep("^YDL248W$", rownames(a))
a <- as.data.frame(logcounts)
a[1,1]
goMapping$`GO:0006810`

```

</br>


### 11. Pearson Correlation for the Gene Counts

```{r}

fdr <- p.adjust(pvalue,"fdr")

```

</br>

### Session Information 

```{r Session Info, echo=FALSE}
sessionInfo()
```