---
title: "Yeast Gene Imputation (blinded data)"
author: "M.Soria"
date: "19/10/2020"
output: html_document
---
</br>

***

##### The following code imputes function on genes from Sacharromyces cerevisiae. Some of the genes were randomly chosen and "blinded" for testing recall.
***

</br>

### Prerequisite libraries
```{r Libraries, echo=TRUE}
suppressPackageStartupMessages({
    library(R.utils)
    library(dplyr)
    library(tidyr)
    library(data.table)
    library(RColorBrewer)
    library(gplots)
    library(Polychrome)
    library(tidyverse)
    library(caret)
    library(ggplot2)
    library(reshape2)
})

source("functions.R")

```
</br>

### Load Pre-requisite RDS files 
```{r echo=TRUE}
file_logcounts_PASS <- "logcounts_PASS.rds"
file_agg <- "agg.rds"
file_clusters <- "clusters.rds"
file_GO <-"GO_table.rds"
file_GO_allClusters <- "GOterms_allClusters.rds"
file_corr_allClusters <- "corr_allClusters.rds"
file_wGO_allClusters <- "wGO_allClusters.rds"
file_GOterms_perCl <- "GOterms_perCl.rds"
file_cuttree_values <- "cuttree_values.rds"

logcounts <- readRDS(file_logcounts_PASS)
agg <- readRDS(file_agg)
clusters <- readRDS(file_clusters)
GO_table <- readRDS(file_GO)
GO_allClusters <-  readRDS(file_GO_allClusters)
corr_allClusters <-  readRDS(file_corr_allClusters)
wGO_allClusters <- readRDS(file_wGO_allClusters)
GOterms_perCl <- readRDS(file_GOterms_perCl)
cuttree_values <- readRDS(file_cuttree_values) 
```

### 1. Blinding known GO terms

```{r}
#Fraction of the total number of genes to be blinded
nfolds <- 10

set.seed(42)
folds <- sample(1:nfolds, nrow(GO_table), replace = TRUE)
#fix rownames
dict_folds <- data.frame(GO_table$GeneID, folds)

saveRDS(dict_folds, "dict_folds.rds")


# Make a data matrix with the blinded genes = 0
which_fold <- 3

ind_blinded <- which(dict_folds$folds == which_fold)
GO_blinded <- GO_table

GO_blinded[ind_blinded, 2:ncol(GO_blinded)] = 0

```

### 3. Impute using blinded GO terms 

```{r }

# Get the GO terms by cluster using the blinded data 
# and the GO list per cluster


GO_blindedClusters <- GO_per_cl_blinded(GO_blinded, 
                        clusters, GOterms_perCl, 100)

wGO_blinded <- impute(cl_GOall = GO_blindedClusters, 
     corr_clAll = corr_allClusters,
     clust_total = 100, thresh = 0.02)

# for 20 clusters
cluster20 <- cuttree_values[["20"]][["GeneID_assignments"]]
corr_allClusters20 <- corr_per_clust(logcounts, cluster20, 20)

GOterms_perCl20 <- GO_per_cl(GO_table, cluster20, 20)
GO_list_perCl20 <- GO_list_perClBlind(GOterms_perCl20, 20)

GO_blindedClusters20 <- GO_per_cl_blinded(GO_blinded, 
                        cluster20, GO_list_perCl20, 20)


wGO_blinded20 <- impute(cl_GOall=GOterms_perCl20, 
                        corr_clAll=corr_allClusters20, 
                        clust_total=20, thresh=0)


```
</br>

### 4. Measures of performance 

```{r}

# per cluster

wGO_allClusters20 <- impute(cl_GOall = GOterms_perCl20, 
     corr_clAll = corr_allClusters20,
     clust_total = 20, thresh = 0)

stats_perCl20 <- stats_cl(wGO_allClusters20, wGO_blinded20, 20)


#for the entire data frame

stats_final20 <- stats_all(stats_perCl20)

# does not work for other R server (R version 3.6.3 (2020-02-29)
# Platform: x86_64-pc-linux-gnu (64-bit))
# stats_df2 <- do.call(rbind.data.frame, stats_perCl)

# works for both servers
stats_df20 <- as.data.frame(t(as.data.frame(lapply(stats_perCl20, unlist))))


```

### 5. K-Fold with Global perfomance metrics (per threshold)

```{r}

start.time <- Sys.time()

kfold <- cross_val(n=10, GO_annot=GO_table, clusters=clusters,
                      GO_list_perCl=GOterms_perCl, 
                      corr_clAll=corr_allClusters,
                      wGO_allClusters=wGO_allClusters,
                      clust_total=100, thresh=0.02)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken # 1.801095 hours

kfold_50 <- cross_val(n=10, GO_annot=GO_table, clusters=clusters,
                      GO_list_perCl=GOterms_perCl, 
                      corr_clAll=corr_allClusters,
                      wGO_allClusters=wGO_allClusters,
                      clust_total=100, thresh=0.5)

kfold20_40 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster20,
                      GO_list_perCl=GO_list_perCl20, 
                      corr_clAll=corr_allClusters20,
                      wGO_allClusters=wGO_allClusters20,
                      clust_total=20, thresh=0.4)

kfold_df <- as.data.frame(do.call(rbind, kfold))
kfold_90_df <- as.data.frame(do.call(rbind, kfold_90))
kfold_50_df <- as.data.frame(do.call(rbind, kfold_50))

kfold_50_ave <- mean(unlist(kfold_50_df$Precision))



cross_val_thresh <- function(interval){
    
    kfold_gobal <- list()
    
    thresh <- seq(from = 0, to = 1, by = interval)
    
    for (i in thresh){
        
        kfold <- cross_val(n=10, GO_annot=GO_table, clusters=cluster20,
                      GO_list_perCl=GOterms_perCl, 
                      corr_clAll=corr_allClusters,
                      wGO_allClusters=wGO_allClusters,
                      clust_total=20, thresh= i)
        
        kfold_df <- as.data.frame(do.call(rbind, kfold))
        
        Sensitivity <- mean(unlist(kfold_df$Sensitivity))
        Specificity <- mean(unlist(kfold_df$Specificity))
        Precision <- mean(unlist(kfold_df$Precision))
        Accuracy <- mean(unlist(kfold_df$Accuracy))
        F1_Score <- mean(unlist(kfold_df$`F1 Score`))
        
        
        kfold_gobal[[paste0(i)]][["DF"]] <- kfold_df
        kfold_gobal[[paste0(i)]][["Sensitivity"]] <- Sensitivity
        kfold_gobal[[paste0(i)]][["Specificity"]] <- Specificity
        kfold_gobal[[paste0(i)]][["Precision"]] <- Precision
        kfold_gobal[[paste0(i)]][["Accuracy"]] <- Accuracy
        kfold_gobal[[paste0(i)]][["F1_Score"]] <- F1_Score
    }
    return(kfold_gobal)
}


kfold_mult <- cross_val_thresh(0.4)
# 5.564071 hours

kfold_mult2 <- cross_val_thresh(0.2)

start.time <- Sys.time()

kfold_mult20 <- cross_val_thresh(0.4)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken # 5.564071 hours
```

### 6. Coarse Sweep

```{r}
# Coarse sweep of the following parameters: cluster total(20, 50, 200, 500) and threshold (0.1, 0.4, 0.8)

#cuttree_values[["32"]][["GeneID_assignments"]]

# ---------------------- for 20 clusters (run inputs)
cluster20 <- cuttree_values[["20"]][["GeneID_assignments"]]
corr_allClusters20 <- corr_per_clust(logcounts, cluster20, 20)
GOterms_perCl20 <- GO_per_cl(GO_table, cluster20, 20)
GO_list_perCl20 <- GO_list_perClBlind(GOterms_perCl20, 20)

GO_blindedClusters20 <- GO_per_cl_blinded(GO_blinded, 
                        cluster20, GO_list_perCl20, 20)

wGO_blinded20 <- impute(cl_GOall=GOterms_perCl20, 
                        corr_clAll=corr_allClusters20, 
                        clust_total=20, thresh=0)

# stats for thresholds 
wGO_allClusters20_10 <- impute(cl_GOall = GOterms_perCl20, 
                        corr_clAll = corr_allClusters20,
                        clust_total = 20, thresh = 0.1)
kfold20_10 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster20,
                       GO_list_perCl=GO_list_perCl20, 
                       corr_clAll=corr_allClusters20,
                       wGO_allClusters=wGO_allClusters20_10,
                       clust_total=20, thresh=0.1)
saveRDS(kfold20_10, "kfold20_10.rds")



wGO_allClusters20_40 <- impute(cl_GOall = GOterms_perCl20, 
                        corr_clAll = corr_allClusters20,
                        clust_total = 20, thresh = 0.4)
kfold20_40 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster20,
                       GO_list_perCl=GO_list_perCl20, 
                       corr_clAll=corr_allClusters20,
                       wGO_allClusters=wGO_allClusters20_40,
                       clust_total=20, thresh=0.4)
saveRDS(kfold20_40, "kfold20_40.rds")



start.time <- Sys.time()
wGO_allClusters20_80 <- impute(cl_GOall = GOterms_perCl20, 
     corr_clAll = corr_allClusters20,
     clust_total = 20, thresh = 0.8)

kfold20_80 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster20,
                       GO_list_perCl=GO_list_perCl20, 
                       corr_clAll=corr_allClusters20,
                       wGO_allClusters=wGO_allClusters20_80,
                       clust_total=20, thresh=0.8)
saveRDS(kfold20_80, "kfold20_80.rds")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken

# ---------------------- for 50 clusters (run inputs)
cluster50 <- cuttree_values[["50"]][["GeneID_assignments"]]
corr_allClusters50 <- corr_per_clust(logcounts, cluster50, 50)
GOterms_perCl50 <- GO_per_cl(GO_table, cluster50, 50)
GO_list_perCl50 <- GO_list_perClBlind(GOterms_perCl50, 50)

GO_blindedClusters50 <- GO_per_cl_blinded(GO_blinded, 
                        cluster50, GO_list_perCl50, 50)

#wGO_blinded50 <- impute(cl_GOall=GOterms_perCl50, 
#                        corr_clAll=corr_allClusters50, 
#                        clust_total=50, thresh=0)

# stats for thresholds 
wGO_allClusters50_10 <- impute(cl_GOall = GOterms_perCl50, 
                        corr_clAll = corr_allClusters50,
                        clust_total = 50, thresh = 0.1)
kfold50_10 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster50,
                       GO_list_perCl=GO_list_perCl50, 
                       corr_clAll=corr_allClusters50,
                       wGO_allClusters=wGO_allClusters50_10,
                       clust_total=50, thresh=0.1)
saveRDS(kfold50_10, "kfold50_10.rds")


wGO_allClusters50_40 <- impute(cl_GOall = GOterms_perCl50, 
                        corr_clAll = corr_allClusters50,
                        clust_total = 50, thresh = 0.4)
kfold50_40 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster50,
                       GO_list_perCl=GO_list_perCl50, 
                       corr_clAll=corr_allClusters50,
                       wGO_allClusters=wGO_allClusters50_40,
                       clust_total=50, thresh=0.4)
saveRDS(kfold50_40, "kfold50_40.rds")

start.time <- Sys.time()
wGO_allClusters50_80 <- impute(cl_GOall = GOterms_perCl50, 
                        corr_clAll = corr_allClusters50,
                        clust_total = 50, thresh = 0.8)

kfold50_80 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster50,
                       GO_list_perCl=GO_list_perCl50, 
                       corr_clAll=corr_allClusters50,
                       wGO_allClusters=wGO_allClusters50_80,
                       clust_total=50, thresh=0.8)
saveRDS(kfold50_80, "kfold50_80.rds")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken


# ---------------------- for 200 clusters (run inputs)
cluster200 <- cuttree_values[["200"]][["GeneID_assignments"]]
corr_allClusters200 <- corr_per_clust(logcounts, cluster200, 200)
GOterms_perCl200 <- GO_per_cl(GO_table, cluster200, 200)
GO_list_perCl200 <- GO_list_perClBlind(GOterms_perCl200, 200)

GO_blindedClusters200 <- GO_per_cl_blinded(GO_blinded, 
                        cluster200, GO_list_perCl200, 200)

# stats for thresholds 
wGO_allClusters200_10 <- impute(cl_GOall = GOterms_perCl200, 
                        corr_clAll = corr_allClusters200,
                        clust_total = 200, thresh = 0.1)

impute <- function (cl_GOall, corr_clAll, clust_total, thresh){
  wGO_list <- list()
  
  for (i in 1:clust_total){
    
    GO_list <- rownames(cl_GOall[[i]])
    GO_list <- GO_list[order(GO_list)]
    
    corr_cl <- corr_clAll[[i]]
    corr_cl <- corr_cl[order(rownames(corr_cl)),]
    corr_cl <- corr_cl[,order(colnames(corr_cl))]
    gene_list <- rownames(corr_cl)
    
    cl_go <- cl_GOall[[i]]
    
    #If all the genes in the cluster have no GOs and the matrix is empty
    if (is_empty(cl_go) == TRUE){
      wGO_list[[paste0("Cluster", i)]][[paste0("Comment")]] <- "No Gene Ontologies found"
      wGO_list[[paste0("Cluster", i)]][[paste0("Input")]] <- 0
      wGO_list[[paste0("Cluster", i)]][[paste0("Output")]] <- 0
      wGO_list[[paste0("Cluster", i)]][[paste0("Diff")]] <- 0
      
      next
    }
    
    # should add gene that are not included in the GO data otherwise merge 
    # will yield weird results
    diff_go_genes <- setdiff(gene_list,rownames(cl_go))
    add <- data.frame(matrix(0,nrow=length(diff_go_genes),ncol=ncol(cl_go)))
    rownames(add) <- diff_go_genes
    colnames(add) <- colnames(cl_go)
    cl_go <- rbind(cl_go,add)
    cl_go <- cl_go[order(rownames(cl_go)),]
    cl_go <- cl_go[,order(colnames(cl_go))]
    
    wGO_cl <- wcorr_cluster(gene_list=gene_list, corr_cl=corr_cl, cl_GO=cl_go)
    wGO_df <- as.data.frame(do.call(rbind, wGO_cl))
    wGO_thresh <- (as.matrix(wGO_df) > thresh)*1 
    wGO_thresh <- wGO_thresh[order(rownames(wGO_thresh)),]
    wGO_thresh <- wGO_thresh[,order(colnames(wGO_thresh))]
    
    cl_subtract <- wGO_thresh - cl_go
    # need to flip some of the values where NAs were replaced with zeros.
    cl_subtract[cl_subtract<0] <- 1
    
    wGO_list[[paste0("Cluster", i)]][[paste0("Input")]] <- cl_go
    wGO_list[[paste0("Cluster", i)]][[paste0("Output")]] <- wGO_thresh
    wGO_list[[paste0("Cluster", i)]][[paste0("Diff")]] <- cl_subtract
    
  }
  return(wGO_list)
}

kfold200_10 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster200,
                       GO_list_perCl=GO_list_perCl200, 
                       corr_clAll=corr_allClusters200,
                       wGO_allClusters=wGO_allClusters200_10,
                       clust_total=200, thresh=0.1)
saveRDS(kfold200_10, "kfold200_10.rds")


stats_cl <- function(input_df=wGO_allClusters, blinded_df=wGO_blinded, clust_total=200){
  stats_list <- list()
  
  for (i in 1:clust_total){
    input <- input_df[[i]][["Input"]]
    blind <- blinded_df[[i]][["Diff"]]
    
    diff <- input - blind 
    sum <- input + blind
    
    TP <- length(which(sum == 2)) #True positive
    TN <- length(which(sum == 0)) #True negative
    FP <- length(which(diff == -1)) #False positive
    FN <- length(which(diff == 1)) #False negative
    
    # False negative rate (FNR)
    FNR <- FN/(FN+TP)
    # Sensitivity/True Positive Rate (TPR)
    TPR <- 1 - FNR
    # False Positive Rate (FPR)
    FPR <- FP/(FP+TN)
    # Specificity/True Negative Rate (TNR)
    TNR <- 1 - FPR
    # Precision/Positive Predictive Value (PPV)
    PPV <- TP/(TP+FP)
    # accuracy (ACC)
    ACC <- (TP+TN)/(TP+TN+FP+FN)
    # F1 score (is the harmonic mean of precision and sensitivity)
    F1 <- (2*TP)/((2*TP)+FP+FN)
    
    stats_list[[paste0("Cluster", i)]][[paste0("TP")]] <- TP
    stats_list[[paste0("Cluster", i)]][[paste0("TN")]] <- TN
    stats_list[[paste0("Cluster", i)]][[paste0("FP")]] <- FP
    stats_list[[paste0("Cluster", i)]][[paste0("FN")]] <- FN
    
    stats_list[[paste0("Cluster", i)]][[paste0("Sensitivity (TPR)")]] <- TPR
    stats_list[[paste0("Cluster", i)]][[paste0("Specificity (TNR)")]] <- TNR
    stats_list[[paste0("Cluster", i)]][[paste0("Precision (PPV)")]] <- PPV
    stats_list[[paste0("Cluster", i)]][[paste0("Accuracy (ACC)")]] <- ACC
    stats_list[[paste0("Cluster", i)]][[paste0("F1_Score")]] <- F1
  }
  return(stats_list)
}


GO_per_cl_blinded <- function(x=wGO_blinded, y=GO_list_perCl, clust_total=200){
  GO_cl <- list()
  for (i in 1:clust_total){
    clusterX <- y[y$ClusterNumber == i,]
    cluster_list <- as.list(clusterX$GeneID)
    cluster_GOterms <- x[x$GeneID %in% cluster_list,]
    rownames(cluster_GOterms)<- cluster_GOterms[,1] 
    cluster_GOterms[,1] <- c()
    cluster_GOterms <- cluster_GOterms[, colnames(cluster_GOterms) %in% GO_list_perCl[[i]] ]
    GO_cl[[paste0("Cluster", i)]] <- cluster_GOterms
  }
  return(GO_cl)
}


wGO_allClusters200_40 <- impute(cl_GOall = GOterms_perCl200, 
                        corr_clAll = corr_allClusters200,
                        clust_total = 200, thresh = 0.4)
kfold200_40 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster200,
                       GO_list_perCl=GO_list_perCl200, 
                       corr_clAll=corr_allClusters200,
                       wGO_allClusters=wGO_allClusters200_40,
                       clust_total=200, thresh=0.4)
saveRDS(kfold200_40, "kfold200_40.rds")

start.time <- Sys.time()
wGO_allClusters200_80 <- impute(cl_GOall = GOterms_perCl200, 
                        corr_clAll = corr_allClusters200,
                        clust_total = 200, thresh = 0.8)

kfold200_80 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster200,
                       GO_list_perCl=GO_list_perCl200, 
                       corr_clAll=corr_allClusters200,
                       wGO_allClusters=wGO_allClusters200_80,
                       clust_total=200, thresh=0.8)
saveRDS(kfold200_80, "kfold200_80.rds")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken

# ---------------------- for 500 clusters (run inputs)
cluster500 <- cuttree_values[["500"]][["GeneID_assignments"]]
corr_allClusters500 <- corr_per_clust(logcounts, cluster500, 500)
GOterms_perCl500 <- GO_per_cl(GO_table, cluster500, 500)
GO_list_perCl500 <- GO_list_perClBlind(GOterms_perCl500, 500)

GO_blindedClusters500 <- GO_per_cl_blinded(GO_blinded, 
                        cluster500, GO_list_perCl500, 500)

# stats for thresholds 
wGO_allClusters500_10 <- impute(cl_GOall = GOterms_perCl500, 
                        corr_clAll = corr_allClusters500,
                        clust_total = 500, thresh = 0.1)
kfold500_10 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster500,
                       GO_list_perCl=GO_list_perCl500, 
                       corr_clAll=corr_allClusters500,
                       wGO_allClusters=wGO_allClusters500_10,
                       clust_total=500, thresh=0.1)
saveRDS(kfold500_10, "kfold500_10.rds")


wGO_allClusters500_40 <- impute(cl_GOall = GOterms_perCl500, 
                        corr_clAll = corr_allClusters500,
                        clust_total = 500, thresh = 0.4)
kfold500_40 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster500,
                       GO_list_perCl=GO_list_perCl500, 
                       corr_clAll=corr_allClusters500,
                       wGO_allClusters=wGO_allClusters500_40,
                       clust_total=500, thresh=0.4)
saveRDS(kfold500_40, "kfold500_40.rds")

start.time <- Sys.time()
wGO_allClusters500_80 <- impute(cl_GOall = GOterms_perCl500, 
                        corr_clAll = corr_allClusters500,
                        clust_total = 500, thresh = 0.8)

kfold500_80 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster500,
                       GO_list_perCl=GO_list_perCl500, 
                       corr_clAll=corr_allClusters500,
                       wGO_allClusters=wGO_allClusters500_80,
                       clust_total=500, thresh=0.8)
saveRDS(kfold500_80, "kfold500_80.rds")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken


# ---------------------- for 300 clusters (run inputs)
cluster300 <- cuttree_values[["300"]][["GeneID_assignments"]]
corr_allClusters300 <- corr_per_clust(logcounts, cluster300, 300)
GOterms_perCl300 <- GO_per_cl(GO_table, cluster300, 300)
GO_list_perCl300 <- GO_list_perClBlind(GOterms_perCl300, 300)

GO_blindedClusters300 <- GO_per_cl_blinded(GO_blinded, 
                        cluster300, GO_list_perCl300, 300)

# stats for thresholds 
wGO_allClusters300_10 <- impute(cl_GOall = GOterms_perCl300, 
                        corr_clAll = corr_allClusters300,
                        clust_total = 300, thresh = 0.1)
kfold300_10 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster300,
                       GO_list_perCl=GO_list_perCl300, 
                       corr_clAll=corr_allClusters300,
                       wGO_allClusters=wGO_allClusters300_10,
                       clust_total=300, thresh=0.1)
saveRDS(kfold300_10, "kfold300_10.rds")


wGO_allClusters300_40 <- impute(cl_GOall = GOterms_perCl300, 
                        corr_clAll = corr_allClusters300,
                        clust_total = 300, thresh = 0.4)
kfold300_40 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster300,
                       GO_list_perCl=GO_list_perCl300, 
                       corr_clAll=corr_allClusters300,
                       wGO_allClusters=wGO_allClusters300_40,
                       clust_total=300, thresh=0.4)
saveRDS(kfold300_40, "kfold300_40.rds")

start.time <- Sys.time()
wGO_allClusters300_80 <- impute(cl_GOall = GOterms_perCl300, 
                        corr_clAll = corr_allClusters300,
                        clust_total = 300, thresh = 0.8)

kfold300_80 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster300,
                       GO_list_perCl=GO_list_perCl300, 
                       corr_clAll=corr_allClusters300,
                       wGO_allClusters=wGO_allClusters300_80,
                       clust_total=300, thresh=0.8)
saveRDS(kfold300_80, "kfold300_80.rds")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken


# ---------------------- for 100 clusters (run inputs)
cluster100 <- cuttree_values[["100"]][["GeneID_assignments"]]
corr_allClusters100 <- corr_per_clust(logcounts, cluster100, 100)
GOterms_perCl100 <- GO_per_cl(GO_table, cluster100, 100)
GO_list_perCl100 <- GO_list_perClBlind(GOterms_perCl100, 100)

GO_blindedClusters100 <- GO_per_cl_blinded(GO_blinded, 
                        cluster100, GO_list_perCl100, 100)

# stats for thresholds 
wGO_allClusters100_10 <- impute(cl_GOall = GOterms_perCl100, 
                        corr_clAll = corr_allClusters100,
                        clust_total = 100, thresh = 0.1)
kfold100_10 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster100,
                       GO_list_perCl=GO_list_perCl100, 
                       corr_clAll=corr_allClusters100,
                       wGO_allClusters=wGO_allClusters100_10,
                       clust_total=100, thresh=0.1)
saveRDS(kfold100_10, "kfold100_10.rds")


wGO_allClusters100_40 <- impute(cl_GOall = GOterms_perCl100, 
                        corr_clAll = corr_allClusters100,
                        clust_total = 100, thresh = 0.4)
kfold100_40 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster100,
                       GO_list_perCl=GO_list_perCl100, 
                       corr_clAll=corr_allClusters100,
                       wGO_allClusters=wGO_allClusters100_40,
                       clust_total=100, thresh=0.4)
saveRDS(kfold100_40, "kfold100_40.rds")

start.time <- Sys.time()
wGO_allClusters100_80 <- impute(cl_GOall = GOterms_perCl100, 
                        corr_clAll = corr_allClusters100,
                        clust_total = 100, thresh = 0.8)

kfold100_80 <- cross_val(n=10, GO_annot=GO_table, clusters=cluster100,
                       GO_list_perCl=GO_list_perCl100, 
                       corr_clAll=corr_allClusters100,
                       wGO_allClusters=wGO_allClusters100_80,
                       clust_total=100, thresh=0.8)
saveRDS(kfold100_80, "kfold100_80.rds")
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken



```

### 7. Visualizations

```{r}
# 
# # correlations of the counts (Pearson - default; Spearman)
# corr_logcounts <- cor(t(logcounts))
# saveRDS(corr_logcounts,"corr_logcounts.rds")
# 
# corr_logcounts_spearman <- cor(t(logcounts), method = "spearman")
# saveRDS(corr_logcounts,"corr_logcounts_spearman.rds")
# 
# # Get lower triangle of the correlation matrix
# corr_lowertri <- corr_logcounts
# corr_lowertri[upper.tri(corr_counts_lowertri)] <- NA
# 
# # Melt the correlation matrix
# melted_cormat <- melt(corr_lowertri, na.rm = TRUE)
# melted_cormat <- melted_cormat[which(melted_cormat$Var1 != melted_cormat$Var2),]
# melted_cormat <- melted_cormat[which(melted_cormat$value != 1),]
# melted_cormat <- melted_cormat[order(-melted_cormat$value),]
# 
# # Heatmap
# png("Data/yeast_heatmaps_PASS_corrlogcounts4.png",    # create PNG for the heat map        
#   width = 5*300,        # 5 x 300 pixels
#   height = 5*300,
#   res = 300,            # 300 pixels per inch
#   pointsize = 8)        # smaller font size
# 
# ggplot(data = melted_cormat[1:10000,], aes(Var2, Var1, fill = value))+
#  geom_tile(color = "white")+
#  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
#    midpoint = 0, limit = c(-1,1), space = "Lab", 
#    name="Pearson\nCorrelation") +
#   theme_minimal()+  
#   theme(axis.title.x=element_blank(),
#         axis.text.x=element_blank(),
#         axis.ticks.x=element_blank()) +
#  coord_fixed()
# 
# dev.off()         
# 
# 
# # creates a own color palette from red to green
# #my_palette <- colorRampPalette(c("yellow", "orange", "red"))
# my_palette <- colorRampPalette(c("red", "black", "green"))
# 
# 
# # (optional) defines the color breaks manually for a "skewed" color transition
# col_breaks = c(seq(-1,0,length=100), # for red
# seq(0.01,0.8,length=100),  # for yellow
# seq(0.81,1,length=100)) # for green
# 
# start.time <- Sys.time()
# png("Data/yeast_heatmaps_PASS_corrlogcounts3.png",    # create PNG for the heat map        
#   width = 5*300,        # 5 x 300 pixels
#   height = 5*300,
#   res = 300,            # 300 pixels per inch
#   pointsize = 8)        # smaller font size
# 
# heatmap.2(melted_cormat[1:10000,],
#   main = "Correlation of Counts", # heat map title
#   margins =c(5,5),     # widens margins around plot
#   col=my_palette,       # use on color palette defined earlier
#   #breaks=col_breaks,    # enable color transition at specified limits
#   trace="none",
#   dendrogram="both",     # only draw a row dendrogram
#   Rowv=T,
#   Colv=T,
#   symm=F,symkey=F,symbreaks=T, scale="none")            
# 
# 
# dev.off()          
# end.time <- Sys.time()
# time.taken <- end.time - start.time
# time.taken
# 
# 
# # histogram - Pearson Method
# png("Data/yeast_histograms_PASS_corrlogcounts.png",    # create PNG for the heat map        
#   width = 5*300,        # 5 x 300 pixels
#   height = 5*300,
#   res = 300,            # 300 pixels per inch
#   pointsize = 8)        # smaller font size
# 
# hist(melted_cormat, 
#      main="Correlation of Counts (Pearson Method)", 
#      xlab="Normalised RNA Seq Counts") 
# 
# dev.off()          
# 
# # histogram - Spearman Method
# png("Data/yeast_histograms_PASS_corrlogcounts_spearman.png",       
#   width = 5*300,        # 5 x 300 pixels
#   height = 5*300,
#   res = 300,            # 300 pixels per inch
#   pointsize = 8)        # smaller font size
# 
# hist(corr_logcounts, 
#      main="Correlation of Counts (Spearman Method)", 
#      xlab="Normalised RNA Seq Counts") 
# 
# dev.off()          
# 
# 
# # cut tree value vs Total Clusters – line diagram  
# cuttree_totals_df <- as.data.frame(as.numeric(names(cuttree_values)))
# m <- list()
# for (i in 1:length(cuttree_values)){
#   x <- cuttree_values[[i]][["Cut_value"]]
#   m[[i]] <- x
# }
# cuttree_totals_df$CuttreeVal <- as.numeric(m)
# colnames(cuttree_totals_df)[1] <- "ClTot"
# 
# png("Data/yeast_line_PASS_cuttree_vs_clTotal.png",       
#   width = 5*300,        # 5 x 300 pixels
#   height = 5*300,
#   res = 300,            # 300 pixels per inch
#   pointsize = 8)        # smaller font size
# 
# ggplot(cuttree_totals_df, aes(x=CuttreeVal, y=ClTot)) +
#   geom_line() + ylab("Total Number of Clusters") + 
#   xlab("Cuttree Values") +
#   ggtitle("Cuttree Value vs Total Number of Clusters") +
#   theme(plot.title = element_text(hjust = 0.5))
# 
# dev.off()   
# 
# 
# # Total genes per cluster  – line diagram  
# clTotals <- do.call(rbind.data.frame, sapply(cuttree_values, "[[", 3))
# clTotals$ClTot <- as.factor(as.numeric(str_extract_all(rownames(clTotals), "^[:digit:]*(?=.)")))
# clTotals$ClNum <- as.numeric(str_extract_all(rownames(clTotals), "(?=.)\\d+$"))
# colnames(clTotals)[1] <- "GeneTot"
# rownames(clTotals) <- c()
# 
# 
# png("Data/yeast_violin_PASS_genesPerCl2.png",       
#   width = 5*300,        # 5 x 300 pixels
#   height = 5*300,
#   res = 300,            # 300 pixels per inch
#   pointsize = 8)        # smaller font size        
# 
# ggplot(clTotals, aes(x=ClNum, y=GeneTot, color=ClTot)) +
#   geom_violin(trim=FALSE) + ylab("Total Number of Genes") + 
#   xlab("Cluster Number") +
#   ggtitle("Total Number of Genes per Cluster") +
#   theme(plot.title = element_text(hjust = 0.5)) +
#   theme(legend.position="none")
# 
# dev.off()   

```

### 8. GO Summary

```{r}

# no_ontologies1 <- setdiff(unique(rownames(agg)),unique(GO_table$GeneID))
# no_ontologies2 <- which(rowSums(GO_table[,2:length(GO_table)]) == 0)
# 
# length(no_ontologies1)
# length(no_ontologies2) 
# 
# 
# summary_sweep2 <- data.frame()
# summary_sweep2 <- do.call(rbind.data.frame, lapply(kfold20_10, "[", 10))
# colnames(summary_sweep)[1] <- "k20_10.F1_Score"
# summary_sweep2$k20_40 <- do.call(rbind.data.frame, lapply(kfold20_40, function(x) {x[length(x)]}))
# summary_sweep$k20_80 <- do.call(rbind.data.frame, lapply(kfold20_80, function(x) {x[length(x)]}))
# summary_sweep$k50_10 <- do.call(rbind.data.frame, lapply(kfold50_10, function(x) {x[length(x)]}))
# summary_sweep$k50_40 <- do.call(rbind.data.frame, lapply(kfold50_40, function(x) {x[length(x)]}))
# summary_sweep$k50_80 <- do.call(rbind.data.frame, lapply(kfold50_80, function(x) {x[length(x)]}))
# summary_sweep$k200_10 <- do.call(rbind.data.frame, lapply(kfold200_10, function(x) {x[length(x)]}))
# summary_sweep$k200_40 <- do.call(rbind.data.frame, lapply(kfold200_40, function(x) {x[length(x)]}))
# summary_sweep$k200_80 <- do.call(rbind.data.frame, lapply(kfold200_80, function(x) {x[length(x)]}))
# summary_sweep$k300_10 <- do.call(rbind.data.frame, lapply(kfold300_10, function(x) {x[length(x)]}))
# summary_sweep$k300_40 <- do.call(rbind.data.frame, lapply(kfold300_40, function(x) {x[length(x)]}))
# summary_sweep$k300_80 <- do.call(rbind.data.frame, lapply(kfold300_80, function(x) {x[length(x)]}))
# summary_sweep$k500_10 <- do.call(rbind.data.frame, lapply(kfold500_10, function(x) {x[length(x)]}))
# summary_sweep$k500_40 <- do.call(rbind.data.frame, lapply(kfold500_40, function(x) {x[length(x)]}))
# summary_sweep$k500_80 <- do.call(rbind.data.frame, lapply(kfold500_80, function(x) {x[length(x)]}))
# 
# Mean.F1_Score <- colMeans(summary_sweep)
# summary_sweep <- rbind(summary_sweep, Mean.F1_Score)
# rownames(summary_sweep)[rownames(summary_sweep) == "11"] <- "Mean.F1_Score"
# saveRDS(summary_sweep, "summary_sweep.rds")
# 
# 
# kfold_list <- list(k20_10=kfold20_10, k20_40=kfold20_40, k20_80=kfold20_80, 
#                 k50_10=kfold50_10, k50_40=kfold50_40, k50_80=kfold50_80, 
#                 k200_10=kfold200_10, k200_40=kfold200_40, k200_80=kfold200_80, 
#                 k300_10=kfold300_10, k300_40=kfold300_40, k300_80=kfold300_80,
#                 k500_10=kfold500_10, k500_40=kfold500_40, k500_80=kfold500_80)
# 
#   
# mean_Csweep <- function(kfold_list, stat_type=10) {
#   mean_df <- data.frame()
#   
#   mean_df <- do.call(rbind.data.frame, lapply(kfold_list[[1]], "[", stat_type))
#   colnames(mean_df)[1] <- names(kfold_list[1])
#   
#   for (i in 2:length(kfold_list)){
#     col_name <- names(kfold_list[i])
#     mean_df[col_name] <- do.call(rbind.data.frame, lapply(kfold_list[[i]], "[", stat_type))
#   }
#   
#   Mean <- colMeans(mean_df)
#   mean_df <- rbind(mean_df, Mean)
#   rownames(mean_df)[rownames(mean_df) == "11"] <- "Mean"
#   
#   return(mean_df)
# }
# 
# 
# summary_Csweep <- function(kfold_list){
#   summary_df <- data.frame()
#   summary <- mean_Csweep(kfold_list, stat_type=6)
#   row_name <- names(kfold_list[[1]][[1]][6])
#   summary_df <- summary[11,]
#   rownames(summary_df)[rownames(summary_df) == "Mean"] <- row_name
#   
#   for (i in 7:length(kfold_list)){
#   summary <- mean_Csweep(kfold_list, stat_type=i)
#   row_name <- names(kfold_list[[1]][[1]][i])
#   summary_df[nrow(summary_df)+1,] <- summary[11,]
#   rownames(summary_df)[rownames(summary_df) == "Mean"] <- row_name
#   }
#   return(summary_df)
# }
# 
# 
# sum <- summary_Csweep(kfold_list)



```


### Session Information 

```{r Session Info, echo=FALSE}
sessionInfo()
```
