---
title: 'Yeast Data Preparation: Filtering, Aggregation, and Normalisation'
author: "M.Soria"
date: "`r Sys.Date()`"
output:
  html_document: null
  toc: yes
  pdf_document: default
theme: cosmo
---

### Introduction

The following code consists of the data preparation steps for the RNA-Seq data for Saccharomyces cerevisiae. The data preparation consists of quality filtering, aggregation, and normalisation.

Clustering of the genes from this counts data for a series of cluster sizes is also performed to be used in the optimasation process.

The GO annotation for yeast is also processed and transformed into a binary matrix for downstream processes.

### Prerequisite libraries

```{r Libraries, echo=TRUE}

suppressPackageStartupMessages({
    library(R.utils)
    library(dplyr)
    library(tidyr)
    library(data.table)
    library(RColorBrewer)
    library(gplots)
    library(edgeR)
    library(reshape2)
    library(Polychrome)
    library(dendextend)
    library(ggplot2)
    library(DESeq2)
    library(dynamicTreeCut)
})

source("functions.R")

```

### 1. Download and unzip the Yeast data 

```{r Download Data, echo=TRUE}

# if statement checks if object exists

if (!file.exists("Data/scerevisiae_se.tsv.bz2")){
    download.file("http://dee2.io/mx/scerevisiae_se.tsv.bz2", 
              destfile = "Data/scerevisiae_se.tsv.bz2")}

if (!file.exists("Data/scerevisiae_se.tsv")){
    bunzip2("Data/scerevisiae_se.tsv.bz2", destname="Data/scerevisiae_se.tsv")}

scerevisiae_data <- fread("Data/scerevisiae_se.tsv", sep = '\t', header = FALSE)

head(scerevisiae_data)

```

### 2. Download, read, and unzip the QC data

```{r QC Data, echo=TRUE}

if (!file.exists("Data/scerevisiae_qc.tsv.bz2")){
    download.file("http://dee2.io/mx/scerevisiae_qc.tsv.bz2", 
              destfile = "Data/scerevisiae_qc.tsv.bz2")}

quality_data <- fread("Data/scerevisiae_qc.tsv.bz2", sep = '\t', 
  col.names = c("Database", "Description", "Details"))

head(quality_data)

```

### 3. Filter and quantify databases according to QC_SUMMARY = PASS, WARN, or FAIL.  

```{r QC Data Summary, echo=TRUE}

# Count the total number of QC_SUMMARY entries for checking
qc_summary <- quality_data[quality_data$Description == 'QC_SUMMARY',]
total_summary <- nrow(qc_summary)

# Count the total number of databases marked as PASS
qc_pass <- filter(quality_data, grepl("PASS", Details))
total_pass <- nrow(qc_pass)

# Count the total number of databases marked as WARN
qc_warn <- filter(quality_data, grepl("WARN", Details))
total_warn <- nrow(qc_warn)

# Count the total number of databases marked as FAIL
qc_fail <- filter(quality_data, grepl("FAIL", Details))
total_fail <- nrow(qc_fail)

# Summary of counts. Addition of each PASS, WARN, and FAIL counts equals to total_summary.
quality_summary <- data.frame(total_pass, total_warn, total_fail, total_summary)
quality_summary

```

### 4. Select all the databses from scerevisiae_se.tsv with QC_SUMMARY == 'PASS' 

```{r Filter yeast bulk data, echo=TRUE}

# Filter all the sample names with QC_SUMMARY == 'PASS' from step 3
databse_pass <- filter(quality_data, grepl("PASS", Details)) 

# Convert the rows into string and store on a list to use for filtering Yeast data
databse_pass_list <- as.list(as.character(databse_pass$Database))

# Filter Yeast data using generated list
scerevisiae_pass <- scerevisiae_data[scerevisiae_data$V1 %in% databse_pass_list,] 

# Convert format from long to wide
scerevisiae_pass_wide <- scerevisiae_pass %>% pivot_wider(names_from = "V1", values_from = "V3")

# Convert tibble to data frame and assign column 1 as rowname
scerevisiae_pass_wide <- as.data.frame(scerevisiae_pass_wide)
rownames(scerevisiae_pass_wide) <- scerevisiae_pass_wide[,1]
scerevisiae_pass_wide <- scerevisiae_pass_wide[,-1]

scerevisiae_pass_wide[1:10,1:6]

```

### 5. Aggregate multiple runs, if any, (SRR) to its corresponding experiment (SRX) with PASS only databases

```{r SRR to SRX Aggregation, echo=TRUE}

# Download and read Metadata summary

if (!file.exists("Data/scerevisiae_metadata.tsv.cut")){
    download.file("http://dee2.io/metadata/scerevisiae_metadata.tsv.cut", 
              destfile = "Data/scerevisiae_metadata.tsv.cut")}
    
yeast_metadata <- read.csv("Data/scerevisiae_metadata.tsv.cut", sep = '\t')

# Filter metadata to include only "passed" samples
yeast_metadata_pass <- yeast_metadata[yeast_metadata$SRR_accession %in% databse_pass_list,]

# Assign column 1 as rownames 
rownames(yeast_metadata_pass) <- yeast_metadata_pass[,1]
yeast_metadata_pass <- yeast_metadata_pass[,-1]

# Put the filtered yeast GeneCount data and the filtered metadata dataframes into a list
yeast_count_metadata <- list(GeneCounts = scerevisiae_pass_wide, MetadataSummary = yeast_metadata_pass)

# Apply both dataframes to the function
agg <- srx_agg(yeast_count_metadata)

# Remove genes with zero total counts 
length(which(rowSums(agg) == 0)) # 83
no_gene_counts <- names(which(rowSums(agg) == 0))

agg <- agg[!(rownames(agg) %in% no_gene_counts),]

# Save agg object
saveRDS(agg, "agg.rds")

# Histogram showing samples (columns) counts
hist(colSums(agg), breaks = 20)

```

### 6. Normalisation of Data (with PASS only databases)

```{r echo=TRUE}

# Filtering low counts
# Check if there are genes with no counts (rowSums==0)
length(which(rowSums(agg) == 0)) # no rowSums = 0

#NORMALISATION FOR COMPOSITION BIAS (TMM normalisation)
dgeObj <- DGEList(agg)
# Apply TMM normalisation to DGEList object
dgeObj <- calcNormFactors(dgeObj, method = "TMM") 
tmm_agg <- cpm(dgeObj)

# Scale after normalisation
tmm_agg_s <- scale(tmm_agg)

# Check distributions of samples
hist(tmm_agg_s)
saveRDS(tmm_agg_s,"tmm_agg_s.rds")

# Correlation Matrix
corr <- cor(t(tmm_agg_s), method = "spearman")
lt <- corr
lt[upper.tri(lt,diag=TRUE)] <- NA

lt_melt <- melt(lt, na.rm = TRUE)

png("Data/hist_vds.png", width = 5*300, height = 5*300, res = 300, pointsize = 8)
hist(lt, breaks = 20, xlab = "Gene Correlation Values", main = "Spearman Correlation of RNASeq Counts")
dev.off()

# Check for 0 values
length(which(lt_melt$value ==0 )) # no zero values


```

### 7a. Hierarchical Clustering: linking method

There are different ways of deciding the distance of elements for them to be grouped into a cluster.The correlation between the distance matrix and the cophenetic distance is calculated to to ensure good
linkage method fit. 

```{r echo=TRUE}
set.seed(42)

cl <-as.dist(1-cor(t(tmm_agg_s), method="spearman"))
hr <- hclust(cl , method="complete")

# Check the correlation between the distance 
# matrix and the cophenetic distance 
cophenetic.d = cophenetic(hr)
corr_cophVSdist <- cor(cl, cophenetic.d)

png("Data/dendro_Average.png", width = 5*300, height = 5*300, res = 300, pointsize = 8)
dend <- as.dendrogram(hr)
labels_colors(dend) <- "white"
plot(dend, main = paste0("Complete Method: Corr = ", signif(corr_cophVSdist)))
dev.off()

```

### 7b. Heirarchical Clustering: deciding the number of clusters
A specific number of clusters can be set using the cutree function.

```{r echo=TRUE}

# Using the argument k in cutree is a faster process vs
# figuring out the cut height
  mycl <- cutree(hr, k=200)
  mycl_length <- length(unique(mycl))
# Cluster size
  mycl_length
# This is the base code used in the cl_lengthCut function below
  clusters <- as.data.frame(mycl)
  colnames(clusters) <- "ClusterNumber"
  clusters$GeneID <- rownames(clusters)
  

```

### 8. Clustering genes using different cluster sizes 

This functions will will yield a list of genes grouped in cluster sized of 50 to 2000 to be used in downstream processes.

```{r eval=FALSE, include=FALSE}

# Different Cluster sizes using the normal dendrogram cutree function
cuttree_values <- cl_lengthCut(hr=hr, min=50, max=2000, interval=2)
saveRDS(cuttree_values, "cuttree_values.rds")

# Different Cluster sizes using the dynamic method, cutreeHybrid from the dynamicTreeCut library
cuttree_values_dynamic <- cl_cut_dynamic(hr=hr, cl=cl, min=4, max=302, interval=2)
saveRDS(cuttree_values_dynamic, "cuttree_values_dynamic.rds")

```


### 9. Heatmaps for the clusters

```{r eval=FALSE, include=FALSE}
# Visualizations
if (!exists("P100")){
  P100 <- createPalette(100,  c("#ff0000", "#00ff00", "#0000ff"))}

clusterCols <- P100[1:100]
colfunc <- colorRampPalette(c("blue", "white", "red"))

# Heatmaps TMM
png("Data/hm_TMM_sp.png", width = 5*300, height = 5*300, res = 300, pointsize = 8)
myClusterSideBar <- clusterCols[tmm_aggCL_sp[["mycl"]]]
heatmap.2(corr, main="Gene Correlation (TMM, Spearman)",  Rowv=as.dendrogram(hr),
          dendrogram="both", scale="col", col = colfunc(25), trace="none",
          RowSideColors= myClusterSideBar, margins = c(5,5))
dev.off()

```

### 10. Gene Ontology and Gene Data 

A binary matrix was constructed with the Entrez Gene IDs (NCBI gene IDs) as rownames and GO IDs as the column names (wide format). If a GO ID is associated with a Gene ID, the cell will equal to 1, otherwise it will be zero. 

```{r}

# Get the GO Annotation File (GAF) format 2.1 of yeast
if (!file.exists("Data/scerevisiae_sgd.gaf.gz")){
    download.file("http://current.geneontology.org/annotations/sgd.gaf.gz", 
              destfile = "Data/scerevisiae_sgd.gaf.gz")}
if (!file.exists("Data/scerevisiae_sgd.gaf")){
    gunzip("Data/scerevisiae_sgd.gaf.gz", destname="Data/scerevisiae_sgd.gaf")}

scerevisiae_GO <- fread("Data/scerevisiae_sgd.gaf", sep = '\t', header = FALSE)

# Subset unique GeneIDs (V11) and GO IDs (V5) in a dataframe. 
# Select the first gene synonym from column V11 and delete the rest. 
# Use this as the GeneID column. 
# Add another column with 1 in each row indicating the relationship of the Gene ID to the GO ID

scerevisiae_GO_matrix <- data.frame(unique(scerevisiae_GO[,c(11,5)]))
scerevisiae_GO_matrix$GeneID <- gsub("\\|.*", "", scerevisiae_GO_matrix$V11)
scerevisiae_GO_matrix$V3 <- as.numeric(paste(1, scerevisiae_GO_matrix$V3))
scerevisiae_GO_matrix <- scerevisiae_GO_matrix[,-1]

# Convert the dataframe to wide format with Gene ID as column 1 and GO ID as column names 
scerevisiae_GO_matrix_wide <- scerevisiae_GO_matrix %>% pivot_wider(names_from = "V5", values_from = "V3")
scerevisiae_GO_matrix_wide <- as.data.frame(scerevisiae_GO_matrix_wide)
colnames(scerevisiae_GO_matrix_wide)[1] <- "GeneID"

# Convert NA to Zero
scerevisiae_GO_matrix_wide[is.na(scerevisiae_GO_matrix_wide)] <- 0

# Save the dataframe to an RDS file
saveRDS(scerevisiae_GO_matrix_wide, "scerevisiae_GO_matrix_wide.rds")

# GeneIDs that are in the gene count data but not in the GO matrix
diff_Count.GO <- setdiff(rownames(agg), scerevisiae_GO_matrix_wide$GeneID)

# GeneIDs that are in the GO matrix but not in the  gene count data
diff_GO.Count <- setdiff(scerevisiae_GO_matrix_wide$GeneID, rownames(agg))

# Remove the GeneIDs that are in the GO matrix but not in the gene count data
IndexRemoved <- which(scerevisiae_GO_matrix_wide$GeneID %in% diff_GO.Count)
GO_table <- scerevisiae_GO_matrix_wide
GO_table <- GO_table[-IndexRemoved,]
# Check if the genes were properly removed 
diff_GO.Count2 <- setdiff(GO_table$GeneID, rownames(agg))
dim(diff_GO.Count2)

# Save the new GO table
saveRDS(GO_table, "GO_table.rds")

```

### 11. Blind GO Terms (90-10 split)

10 per cent of GeneIDs will have zero Gene Ontology Annotations. This will serve as the training data. After the model has been optimised, these IDs will be unblided as the testing data.

```{r echo=TRUE}
#Fraction of the total number of genes to be blinded
test_size <- floor(0.10 * nrow(GO_table))

## set the seed to make your partition reproducible
set.seed(42)
blinded_ind <- sample(seq_len(nrow(GO_table)), size = test_size)
GO_train <- GO_table
GO_train[blinded_ind, 2:ncol(GO_train)] = 0

saveRDS(GO_train, "GO_train.rds")

GO_test <- GO_table[blinded_ind,]
saveRDS(GO_test, "GO_test.rds")

tmm_agg_s_test <- tmm_agg_s[rownames(tmm_agg_s) %in% GO_test$GeneID,]
saveRDS(tmm_agg_s_test, "tmm_agg_s_test.rds")
```


### Session Information 

```{r Session Info, echo=FALSE}

sessionInfo()

```
