---
title: "Yeast Gene Imputation (blinded data)"
author: "M.Soria"
date: "19/10/2020"
output: html_document
---
</br>

***

##### The following code imputes function on genes from Sacharromyces cerevisiae. Some of the genes were randomly chosen and "blinded" for testing recall.
***

</br>

### Prerequisite libraries
```{r Libraries, echo=TRUE}
suppressPackageStartupMessages({
    library(R.utils)
    library(dplyr)
    library(tidyr)
    library(data.table)
    library(RColorBrewer)
    library(gplots)
    library(Polychrome)
    library(tidyverse)
    library(caret)
    library(ggplot2)
    library(reshape2)
    library(tools)
    library (igraph)
})

source("functions.R")

```
</br>

### Load Pre-requisite RDS files 
```{r echo=TRUE}
file_PASS <- "tmm_agg_s.rds"
file_tmm_agg_s_test <- "tmm_agg_s_test.rds"
file_GO <-"GO_table.rds"
file_GO_train <- "GO_train.rds"
file_GO_test <- "GO_test.rds"
file_cuttree_values <- "cuttree_values_dynamic.rds"

# Normalised, aggregated, and scaled RNASeq counts
tmm_agg_s <- readRDS(file_PASS)
tmm_agg_s_test <- readRDS(file_tmm_agg_s_test)
# Binary df for GO annotations 
GO_table <- readRDS(file_GO)
# From GO_table, selected randomly as training data
GO_train <- readRDS(file_GO_train)
GO_test <- readRDS(file_GO_test)
# List of genes grouped in different cluster sizes
cuttree_values <- readRDS(file_cuttree_values)
```

### 1. Coarse Sweep for Spearman + complete linkage 
The following code will perform a K-fold validation for the imputation process which uses Spearman method to measure the distance matrix and complete as linkage method during the clustering process. 

```{r eval=FALSE, include=FALSE}
# Coarse sweep of the following parameters: cluster total(304,139,93,54,30) and thresholds (0.1 to 0.9). The comments next to the cluster size in cl_list is the minumum cluster size used in the dynamic tree cut. These values are saved in the cuttree_values variable. 

cl_list <- c(304, #10
             139, #20
             93, #30
             54, #50
             30) #102

#cl_list <- c(93)
thresh_list <- c(0, 1)

sample_kfold <- optimise_impute(cl_list, thresh_list, cuttree_values, tmm_agg_s, GO_train)

saveRDS(sample_kfold, "kfold_dyn_sum01.rds")

```

### 2. dendrogram

```{r}
cl <- readRDS("cl_dist_matrix.rds")
hr <- hclust(cl , method="complete") 
saveRDS(hr,"hr.rds")
    
dend <- as.dendrogram(hr)
c <- cutree(dend, 10)

colored_dend <- color_labels(dend, col = dynamicColors, clusters = as.numeric(dynamicColors), groupLabels = TRUE)
plot(colored_dend)

pdf("Data/dend.pdf", width=80, height=15)
par(cex = 0.05)
plot(dend, main="Gene-to-Gene Correlation Dendrogram of Saccharomyces cerevisiae")
dev.off()

# cluster 3
cl3_139_dend <- find_dendrogram(dend, cl3_139GeneIDS)
colored_139 <- color_labels(cl3_139_dend,labels = "ICR1",col = "red", groupLabels = TRUE)

sub_139 <- c("ICR1", "YNL025C", "YLL006W", "YCL004W", "YDR314C", "YMR154C", "YBR223C", "YHR159W")
sub_139_dend <- find_dendrogram(dend, sub_139)
sub_139_colored <- color_labels(sub_139_dend,labels = "ICR1", col = "blue")

png("Data/sub_cl3_139.png", width = 7*300, height = 4*300, res = 300, pointsize = 8)
par(cex = 1.5)
plot(sub_139_colored)
dev.off()

png("Data/cl3_139.png", width = 7*300, height = 4*300, res = 300, pointsize = 8)
par(cex = 0.8)
plot(colored_139)
dev.off()

# cluster 16
cl16_139_dend <- find_dendrogram(dend, pseudogene_cl14_genes)
colored_139 <- color_labels(cl16_139_dend,labels = "YPL275W",col = "red")

#png("Data/cl3_139.png", width = 7*300, height = 4*300, res = 300, pointsize = 8)
par(cex = 0.8)
plot(colored_139)
#dev.off()


# cluster 1
cl1_139_dend <- find_dendrogram(dend, cl1_geneList)
colored_cl1_139 <- color_labels(cl1_139_dend,labels = "HRA1",col = "red")

par(mfrow=c(3,1))

plot(dend, main="Main")
plot(cut(dend, h=75)$upper, 
     main="Upper tree of cut at h=75")
plot(cut(dend, h=75)$lower[[2]], 
     main="Second branch of lower tree with cut at h=75")


dyn_tree <- cutreeHybrid(
      # Input data: basic tree cutiing
      dendro=hr, distM=as.matrix(cl),
      # Branch cut criteria and options
      minClusterSize = 20, deepSplit = 4,
      # PAM stage options
      pamStage = TRUE, pamRespectsDendro = FALSE,
      respectSmallClusters = TRUE,
      # Various options
      verbose = 1, indent = 0)


dynamicColors = labels2colors(dyn_tree$labels)
table(dynamicColors)
# Plot the dendrogram and colors underneath
pdf("Data/geneTreeLabels.pdf")
plotDendroAndColors(hr, dynamicColors, "Dynamic Tree Cut",
dendroLabels = FALSE, hang = 0.03,
#groupLabels = hr$labels,
addGuide = TRUE,
guideAll = TRUE, guideHang = 0.05,
addTextGuide = TRUE,
main = "Gene dendrogram and module colors")
dev.off()

```

### interactive dendro

```{r}
library()
if (!file.exists("Data/lncRNAs.tsv")){
    download.file("https://www.genenames.org/cgi-bin/genegroup/download?id=788&type=branch", 
              destfile = "Data/lncRNAs.tsv")}
lnc <- fread("Data/lncRNAs.tsv", sep = '\t', header = TRUE)

a <- setdiff(rownames(tmm_agg_s), lnc$`Approved symbol`)

rn
  
```


### 2. Summary of performace measures for Pearson + complete linkage

```{r}

dat_list <- lapply(sample_kfold, function(x) x[names(x)!="Index_folds"])

# List the all the mean predicton scores for each parameter 
summary_all <- t(summary_Csweep(dat_list))
saveRDS(summary_all, "summary_all_dynamic_max.rds")

```

### 3. Visualizations

```{r}

# cut tree value vs Total Clusters – line diagram
cuttree_totals_df <- function(cuttree_values){
  df <- as.data.frame(as.numeric(names(cuttree_values)))
  m <- list()
  for (i in 1:length(cuttree_values)){
    x <- cuttree_values[[i]][["Cut_value"]]
    m[[i]] <- x
  }
  df$CuttreeVal <- as.numeric(m)
  colnames(df)[1] <- "ClTot"
  png("Data/cuttreeVSclTotal.png",
    width = 5*300,        # 5 x 300 pixels
    height = 5*300,
    res = 300,            # 300 pixels per inch
    pointsize = 8)        # smaller font size
  
  ggplot(df, aes(x=CuttreeVal, y=ClTot)) +
    geom_line() + ylab("Total Number of Clusters") +
    xlab("Cuttree Values") +
    ggtitle("Cuttree Value vs Total Number of Clusters") +
    theme(plot.title = element_text(hjust = 0.5))
  
  dev.off()
}

cuttree_totals_df(cuttree_values)

# Total genes per cluster  – violin plot
c_violin <- function(cuttree_values){
  
  clTotals <- do.call(rbind.data.frame, sapply(cuttree_values, "[[", 3))
  clTotals$ClusterSize <- as.factor(as.numeric(str_extract_all(rownames(clTotals), "^[:digit:]*(?=.)")))
  clTotals$ClusterID <- as.numeric(str_extract_all(rownames(clTotals), "(?=.)\\d+$"))
  colnames(clTotals)[1] <- "TotalGenes"
  rownames(clTotals) <- c()
  
  # Filter according to cluster size
  cluster20_100 <- clTotals[clTotals$ClusterSize %in% c(20,50,100),]
  cluster200_500 <- clTotals[clTotals$ClusterSize %in% c(200, 300, 500),]
  cluster800_1500 <- clTotals[clTotals$ClusterSize %in% c(800, 1000, 1500),]
  
  pdf("Data/Violinplots.pdf")
  
  ggplot(cluster20_100, aes(x=ClusterSize, y=TotalGenes, fill=ClusterSize)) +
    geom_violin(trim=FALSE) + ylab("Total Number of Genes") +
    xlab("Cluster Size") +
    ggtitle("Distibution of genes per cluster size (20 to 100)") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_boxplot(width=0.1, fill="white") + theme_minimal() +
    scale_fill_brewer(palette="Blues")
    
  ggplot(cluster200_500, aes(x=ClusterSize, y=TotalGenes, fill=ClusterSize)) +
    geom_violin(trim=FALSE) + ylab("Total Number of Genes") +
    xlab("Cluster Size") +
    ggtitle("Distibution of genes per cluster size (200 to 500)") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_boxplot(width=0.1, fill="white") + theme_minimal() +
    scale_fill_brewer(palette="Dark2")
  
  ggplot(cluster800_1500, aes(x=ClusterSize, y=TotalGenes, fill=ClusterSize)) +
    geom_violin(trim=FALSE) + ylab("Total Number of Genes") +
    xlab("Cluster Size") +
    ggtitle("Distibution of genes per cluster size (800 to 1500)") +
    theme(plot.title = element_text(hjust = 0.5)) +
     geom_boxplot(width=0.1, fill="white") + theme_minimal() + 
    scale_fill_brewer(palette="RdBu")
  
  dev.off()
}

c_violin(cuttree_values) 


```

### Session Information 

```{r Session Info, echo=FALSE}
sessionInfo()
```
