---
title: "Yeast Gene Imputation (blinded data)"
author: "M.Soria"
date: "19/10/2020"
output: html_document
---
</br>

***

##### The following code imputes function on genes from Sacharromyces cerevisiae. Some of the genes were randomly chosen and "blinded" for testing recall.
***

</br>

### Prerequisite libraries
```{r Libraries, echo=TRUE}
suppressPackageStartupMessages({
    library(R.utils)
    library(dplyr)
    library(tidyr)
    library(data.table)
    library(RColorBrewer)
    library(gplots)
    library(Polychrome)
    library(tidyverse)
    library(caret)
    library(ggplot2)
    library(reshape2)
    library(tools)
    library (igraph)
})

source("functions.R")

```
</br>

### Load Pre-requisite RDS files 
```{r echo=TRUE}
file_PASS <- "tmm_agg_s.rds"
file_GO <-"GO_table.rds"
file_GO_train <- "GO_train.rds"
file_cuttree_values <- "cuttree_values_ward.rds"

tmm_agg_s <- readRDS(file_PASS)
GO_table <- readRDS(file_GO)
GO_train <- readRDS(file_GO_train)
cuttree_values <- readRDS(file_cuttree_values)
```

### 1. Coarse Sweep for Spearman + complete linkage (SandC)
The following code will perform a K-fold validation for the imputation process which uses Pearson method to measure the distance matrix and complete as linkage method during the clustering process. 

```{r eval=FALSE, include=FALSE}
# Coarse sweep of the following parameters: cluster total(20, 50, 200, 500) and threshold (0.1, 0.4, 0.8)

cl_list <- c(1000)
thresh_list <- c(0.1,0.5,0.9)

start <- Sys.time()
sample_kfold <- kfold(cl_list, thresh_list, cuttree_values, tmm_agg_s, GO_train)
end <- Sys.time()
totalTime <- end -start
```

### 3. Summary of performace measures for Pearson + complete linkage

```{r}

dat_list <- lapply(CS_ward, function(x) x[names(x)!="Index_folds"])

# List the all the mean predicton scores for each parameter 
summary_all <- summary_Csweep(dat_list)
saveRDS(summary_all, "summary_all.rds")

```

### 4. Visualizations

```{r}

# cut tree value vs Total Clusters – line diagram
cuttree_totals_df <- function(cuttree_values){
  df <- as.data.frame(as.numeric(names(cuttree_values)))
  m <- list()
  for (i in 1:length(cuttree_values)){
    x <- cuttree_values[[i]][["Cut_value"]]
    m[[i]] <- x
  }
  df$CuttreeVal <- as.numeric(m)
  colnames(df)[1] <- "ClTot"
  png("Data/cuttreeVSclTotal.png",
    width = 5*300,        # 5 x 300 pixels
    height = 5*300,
    res = 300,            # 300 pixels per inch
    pointsize = 8)        # smaller font size
  
  ggplot(df, aes(x=CuttreeVal, y=ClTot)) +
    geom_line() + ylab("Total Number of Clusters") +
    xlab("Cuttree Values") +
    ggtitle("Cuttree Value vs Total Number of Clusters") +
    theme(plot.title = element_text(hjust = 0.5))
  
  dev.off()
}

cuttree_totals_df(cuttree_values)

 


# Total genes per cluster  – violin plot
c_violin <- function(cuttree_values){
  
  clTotals <- do.call(rbind.data.frame, sapply(cuttree_values, "[[", 3))
  clTotals$ClusterSize <- as.factor(as.numeric(str_extract_all(rownames(clTotals), "^[:digit:]*(?=.)")))
  clTotals$ClusterID <- as.numeric(str_extract_all(rownames(clTotals), "(?=.)\\d+$"))
  colnames(clTotals)[1] <- "TotalGenes"
  rownames(clTotals) <- c()
  
  # Filter according to cluster size
  cluster20_100 <- clTotals[clTotals$ClusterSize %in% c(20,50,100),]
  cluster200_500 <- clTotals[clTotals$ClusterSize %in% c(200, 300, 500),]
  cluster800_1500 <- clTotals[clTotals$ClusterSize %in% c(800, 1000, 1500),]
  
  pdf("Data/Violinplots.pdf")
  
  ggplot(cluster20_100, aes(x=ClusterSize, y=TotalGenes, fill=ClusterSize)) +
    geom_violin(trim=FALSE) + ylab("Total Number of Genes") +
    xlab("Cluster Size") +
    ggtitle("Distibution of genes per cluster size (20 to 100)") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_boxplot(width=0.1, fill="white") + theme_minimal() +
    scale_fill_brewer(palette="Blues")
    
  ggplot(cluster200_500, aes(x=ClusterSize, y=TotalGenes, fill=ClusterSize)) +
    geom_violin(trim=FALSE) + ylab("Total Number of Genes") +
    xlab("Cluster Size") +
    ggtitle("Distibution of genes per cluster size (200 to 500)") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_boxplot(width=0.1, fill="white") + theme_minimal() +
    scale_fill_brewer(palette="Dark2")
  
  ggplot(cluster800_1500, aes(x=ClusterSize, y=TotalGenes, fill=ClusterSize)) +
    geom_violin(trim=FALSE) + ylab("Total Number of Genes") +
    xlab("Cluster Size") +
    ggtitle("Distibution of genes per cluster size (800 to 1500)") +
    theme(plot.title = element_text(hjust = 0.5)) +
     geom_boxplot(width=0.1, fill="white") + theme_minimal() + 
    scale_fill_brewer(palette="RdBu")
  
  dev.off()
}

c_violin(cuttree_values) 


```

### Session Information 

```{r Session Info, echo=FALSE}
sessionInfo()
```
