</br>
<center> <h1> Yeast Sample Quality Filter </h1> </center>
***

#### The following code will determine the quality of the runs for S.Cerevisiae (Yeast) RNA-seq data downloaded from the [Digital Expression Explorer 2](http://dee2.io/) repository. It will quantify the number of PASS, WARN and FAIL runs and will filter samples with PASS runs  
</br></br></br>


Prerequisite libraries
```{r Libraries, echo=TRUE}
library("R.utils")
library("dplyr")
library("getDEE2")
```


1. Download and unzip the Yeast data 
```{r Download Data, eval=FALSE, include=FALSE}
download.file("http://dee2.io/mx/scerevisiae_se.tsv.bz2", 
              destfile = "Data/scerevisiae_se.tsv.bz2")

R.utils::bunzip2("Data/scerevisiae_se.tsv.bz2", "Data/scerevisiae_se.tsv", overwrite=TRUE, remove=FALSE)
```

2. Download and unzip the QC data
```{r}
download.file("http://dee2.io/mx/scerevisiae_qc.tsv.bz2", 
              destfile = "Data/scerevisiae_qc.tsv.bz2")

R.utils::bunzip2("Data/scerevisiae_qc.tsv.bz2", "Data/scerevisiae_qc.tsv", overwrite=TRUE, remove=FALSE)
```

3. Read Quality data, filter and quantify samples according to QC_SUMMARY = PASS, WARN, or FAIL.  
```{r}
quality_data <- read.csv("Data/scerevisiae_qc.tsv", sep = '\t', col.names = c("Sample", "Description", "Details"))

# count the total number of QC_SUMMARY entries for checking
qc_summary <- quality_data[quality_data$Description == 'QC_SUMMARY',]
total_summary <- nrow(qc_summary)

# count the total number of samples marked as PASS
qc_pass <- filter(quality_data, grepl("PASS", Details))
total_pass <- nrow(qc_pass)

# count the total number of samples marked as WARN
qc_warn <- filter(quality_data, grepl("WARN", Details))
total_warn <- nrow(qc_warn)

# count the total number of samples marked as FAIL
qc_fail <- filter(quality_data, grepl("FAIL", Details))
total_fail <- nrow(qc_fail)

# Summary of counts. Addition of each PASS, WARN, and FAIL counts equals to total_summary.
quality_summary <- data.frame(total_pass, total_warn, total_fail, total_summary)
quality_summary

```

4. Select all the samples from scerevisiae_se.tsv with QC_SUMMARY == 'PASS' 
```{r echo=TRUE}

# filter all the sample names with QC_SUMMARY == 'PASS' from step 3
sample_pass <- filter(quality_data, grepl("PASS", Details)) %>% select(Sample)

# convert the rows into string and store on a list to use for filtering Yeast data
sample_pass_list <- as.list(as.character(sample_pass$Sample))

# read Yeast data
scerevisiae_data <- read.csv("Data/scerevisiae_se.tsv", sep = '\t', col.names = c("Sample", "Gene", "Count"))

# Filter Yeast data using generated list
scerevisiae_pass <- scerevisiae_data[scerevisiae_data$Sample %in% sample_pass_list,]
```

5. Aggregate the SRR run data to SRX experiment 
```{r}
srx_agg <- function(x,counts="GeneCounts") {
    IDX=which(names(x) %in% "GeneCounts")
    mds<-x$MetadataSummary
    n=nrow(x[[IDX]])
    SRX_dat <- vapply(X=unique(mds$SRX_accession) ,
        function(srx) {
            srrs<-rownames(mds)[which(mds$SRX_accession %in% srx)]
        rowSums(x[[IDX]][,srrs])
    } , numeric(n))
    SRX_dat
}

x <- getDEE2("ecoli",c("SRR1613487","SRR1613488"),legacy=TRUE)

# Error 'x' must be an array of at least two dimensions
srx_agg(x, counts="GeneCounts")
```


