---
title: 'Yeast Data Preparation: Filtering, Aggregation, and Normalisation'
author: "M.Soria"
date: "`r Sys.Date()`"
output:
  html_document: null
  toc: yes
  pdf_document: default
theme: cosmo
---

### Introduction

The following code consists of the data preparation steps for the RNA-Seq data for Saccharomyces cerevisiae. 
The data preparation consists of quality filtering, aggregation, and normalisation.

### Prerequisite libraries

```{r Libraries, echo=TRUE}

suppressPackageStartupMessages({
    library(R.utils)
    library(dplyr)
    library(tidyr)
    library(data.table)
    library(RColorBrewer)
    library(gplots)
    library(edgeR)
    library(reshape2)
    library(Polychrome)
    library(dendextend)
    library(ggplot2)
    library(DESeq2)
})

source("functions.R")

```

### 1. Download and unzip the Yeast data 

```{r Download Data, echo=TRUE}

# if statement checks if object exists

if (!file.exists("Data/scerevisiae_se.tsv.bz2")){
    download.file("http://dee2.io/mx/scerevisiae_se.tsv.bz2", 
              destfile = "Data/scerevisiae_se.tsv.bz2")}

if (!file.exists("Data/scerevisiae_se.tsv")){
    bunzip2("Data/scerevisiae_se.tsv.bz2", destname="Data/scerevisiae_se.tsv")}

scerevisiae_data <- fread("Data/scerevisiae_se.tsv", sep = '\t', header = FALSE)

head(scerevisiae_data)

```

### 2. Download, read, and unzip the QC data

```{r QC Data, echo=TRUE}

if (!file.exists("Data/scerevisiae_qc.tsv.bz2")){
    download.file("http://dee2.io/mx/scerevisiae_qc.tsv.bz2", 
              destfile = "Data/scerevisiae_qc.tsv.bz2")}

quality_data <- fread("Data/scerevisiae_qc.tsv.bz2", sep = '\t', 
  col.names = c("Database", "Description", "Details"))

head(quality_data)

```

### 3. Filter and quantify databases according to QC_SUMMARY = PASS, WARN, or FAIL.  

```{r QC Data Summary, echo=TRUE}

# Count the total number of QC_SUMMARY entries for checking
qc_summary <- quality_data[quality_data$Description == 'QC_SUMMARY',]
total_summary <- nrow(qc_summary)

# Count the total number of databases marked as PASS
qc_pass <- filter(quality_data, grepl("PASS", Details))
total_pass <- nrow(qc_pass)

# Count the total number of databases marked as WARN
qc_warn <- filter(quality_data, grepl("WARN", Details))
total_warn <- nrow(qc_warn)

# Count the total number of databases marked as FAIL
qc_fail <- filter(quality_data, grepl("FAIL", Details))
total_fail <- nrow(qc_fail)

# Summary of counts. Addition of each PASS, WARN, and FAIL counts equals to total_summary.
quality_summary <- data.frame(total_pass, total_warn, total_fail, total_summary)
quality_summary

```

### 4. Select all the databses from scerevisiae_se.tsv with QC_SUMMARY == 'PASS' 

```{r Filter yeast bulk data, echo=TRUE}

# Filter all the sample names with QC_SUMMARY == 'PASS' from step 3
databse_pass <- filter(quality_data, grepl("PASS", Details)) 

# Convert the rows into string and store on a list to use for filtering Yeast data
databse_pass_list <- as.list(as.character(databse_pass$Database))

# Filter Yeast data using generated list
scerevisiae_pass <- scerevisiae_data[scerevisiae_data$V1 %in% databse_pass_list,] 

# Convert format from long to wide
scerevisiae_pass_wide <- scerevisiae_pass %>% pivot_wider(names_from = "V1", values_from = "V3")

# Convert tibble to data frame and assign column 1 as rowname
scerevisiae_pass_wide <- as.data.frame(scerevisiae_pass_wide)
rownames(scerevisiae_pass_wide) <- scerevisiae_pass_wide[,1]
scerevisiae_pass_wide <- scerevisiae_pass_wide[,-1]

scerevisiae_pass_wide[1:10,1:6]

```

### 5. Aggregate multiple runs, if any, (SRR) to its corresponding experiment (SRX) with PASS only databases

```{r SRR to SRX Aggregation, echo=TRUE}

# Download and read Metadata summary

if (!file.exists("Data/scerevisiae_metadata.tsv.cut")){
    download.file("http://dee2.io/metadata/scerevisiae_metadata.tsv.cut", 
              destfile = "Data/scerevisiae_metadata.tsv.cut")}
    
yeast_metadata <- read.csv("Data/scerevisiae_metadata.tsv.cut", sep = '\t')

# Filter metadata to include only "passed" samples
yeast_metadata_pass <- yeast_metadata[yeast_metadata$SRR_accession %in% databse_pass_list,]

# Assign column 1 as rownames 
rownames(yeast_metadata_pass) <- yeast_metadata_pass[,1]
yeast_metadata_pass <- yeast_metadata_pass[,-1]

# Put the filtered yeast GeneCount data and the filtered metadata dataframes into a list
yeast_count_metadata <- list(GeneCounts = scerevisiae_pass_wide, MetadataSummary = yeast_metadata_pass)

# Apply both dataframes to the function
agg <- srx_agg(yeast_count_metadata)

# Remove genes with zero total counts 
length(which(rowSums(agg) == 0)) # 83
no_gene_counts <- names(which(rowSums(agg) == 0))

agg <- agg[!(rownames(agg) %in% no_gene_counts),]

# Save agg object
saveRDS(agg, "agg.rds")

# Histogram showing samples (columns) counts
hist(colSums(agg), breaks = 20)

```

### 6. Normalisation of Data (with PASS only databases)

```{r echo=TRUE}

# Filtering low counts
# Check if there are genes with no counts (rowSums==0)
length(which(rowSums(agg) == 0)) # no rowSums = 0

#NORMALISATION FOR COMPOSITION BIAS (TMM normalisation)
dgeObj <- DGEList(agg)
# Apply TMM normalisation to DGEList object
dgeObj <- calcNormFactors(dgeObj, method = "TMM") 
tmm_agg <- cpm(dgeObj)

# Scale after normalisation
tmm_agg_s <- scale(tmm_agg)

# Check distributions of samples
hist(tmm_agg_s)
saveRDS(tmm_agg_s,"tmm_agg_s.rds")

# Correlation Matrix
corr <- cor(t(tmm_agg_s), method = "spearman")
lt <- corr
lt[upper.tri(lt,diag=TRUE)] <- NA

lt_melt <- melt(lt, na.rm = TRUE)

png("Data/hist_vds.png", width = 5*300, height = 5*300, res = 300, pointsize = 8)
hist(lt, breaks = 20, xlab = "Gene Correlation Values", main = "Spearman Correlation of RNASeq Counts")
dev.off()

# Check for 0 values
length(which(lt_melt$value ==0 )) # no zero values


```

### Unsupervised Clustering

```{r echo=TRUE}
set.seed(42)

cl <-as.dist(1-cor(t(tmm_agg_s), method="spearman"))

hr <- hclust(cl , method="ward.D2")
cophenetic.d = cophenetic(hr)
corr_cophVSdist <- cor(corr, cophenetic.d)

png("Data/dendro_ward.png", width = 5*300, height = 5*300, res = 300, pointsize = 8)
avg_dend_obj6 <- as.dendrogram(hr)
labels_colors(avg_dend_obj6) <- "white"
plot(avg_dend_obj6, main = paste0("Ward Method: Corr = ", signif(corr_cophVSdist)))
dev.off()


```

### 7. Heirarchical Clustering with PASS only databases

This chunk can be reused for the PASS + Warn or the WARN filters by changing the content of the filename object. Do not forget to change the name of the variable where you will save the clusters for different inputs. 

```{r echo=TRUE}

  set.seed(42)
  c<-as.dist(1-cor(t(tmm_agg_s), method="spearman"))
  hr <- hclust(c , method="ward.D2")
  mycl <- cutree(hr, k=200)
  mycl_length <- length(unique(mycl))
  mycl_length
  clusters <- as.data.frame(mycl)
  colnames(clusters) <- "ClusterNumber"
  clusters$GeneID <- rownames(clusters)
  saveRDS(clusters, "clusters.rds")

```

### 8. Heatmaps for the clusters

```{r eval=FALSE, include=FALSE}
# Visualizations
if (!exists("P100")){
  P100 <- createPalette(100,  c("#ff0000", "#00ff00", "#0000ff"))}

clusterCols <- P100[1:100]
colfunc <- colorRampPalette(c("blue", "white", "red"))

# Heatmaps TMM
png("Data/hm_TMM_sp.png", width = 5*300, height = 5*300, res = 300, pointsize = 8)
myClusterSideBar <- clusterCols[tmm_aggCL_sp[["mycl"]]]
heatmap.2(corr, main="Gene Correlation (TMM, Spearman)",  Rowv=as.dendrogram(hr),
          dendrogram="both", scale="col", col = colfunc(25), trace="none",
          RowSideColors= myClusterSideBar, margins = c(5,5))
dev.off()

```


### 9. Gene Ontology and Gene Data 

```{r}

# Get the GO Annotation File (GAF) format 2.1 of yeast
if (!file.exists("Data/scerevisiae_sgd.gaf.gz")){
    download.file("http://current.geneontology.org/annotations/sgd.gaf.gz", 
              destfile = "Data/scerevisiae_sgd.gaf.gz")}
if (!file.exists("Data/scerevisiae_sgd.gaf")){
    gunzip("Data/scerevisiae_sgd.gaf.gz", destname="Data/scerevisiae_sgd.gaf")}

scerevisiae_GO <- fread("Data/scerevisiae_sgd.gaf", sep = '\t', header = FALSE)

# Subset unique GeneIDs (V11) and GO IDs (V5) in a dataframe. 
# Select the first gene synonym from column V11 and delete the rest. 
# Use this as the GeneID column. 
# Add another column with 1 in each row indicating the relationship of the Gene ID to the GO ID

scerevisiae_GO_matrix <- data.frame(unique(scerevisiae_GO[,c(11,5)]))
scerevisiae_GO_matrix$GeneID <- gsub("\\|.*", "", scerevisiae_GO_matrix$V11)
scerevisiae_GO_matrix$V3 <- as.numeric(paste(1, scerevisiae_GO_matrix$V3))
scerevisiae_GO_matrix <- scerevisiae_GO_matrix[,-1]

# Convert the dataframe to wide format with Gene ID as column 1 and GO ID as column names 
scerevisiae_GO_matrix_wide <- scerevisiae_GO_matrix %>% pivot_wider(names_from = "V5", values_from = "V3")
scerevisiae_GO_matrix_wide <- as.data.frame(scerevisiae_GO_matrix_wide)
colnames(scerevisiae_GO_matrix_wide)[1] <- "GeneID"

# Convert NA to Zero
scerevisiae_GO_matrix_wide[is.na(scerevisiae_GO_matrix_wide)] <- 0

# Save the dataframe to an RDS file
saveRDS(scerevisiae_GO_matrix_wide, "scerevisiae_GO_matrix_wide.rds")

# GeneIDs that are in the gene count data but not in the GO matrix
diff_Count.GO <- setdiff(rownames(agg), scerevisiae_GO_matrix_wide$GeneID)

# GeneIDs that are in the GO matrix but not in the  gene count data
diff_GO.Count <- setdiff(scerevisiae_GO_matrix_wide$GeneID, rownames(agg))

# Remove the GeneIDs that are in the GO matrix but not in the gene count data
IndexRemoved <- which(scerevisiae_GO_matrix_wide$GeneID %in% diff_GO.Count)
GO_table <- scerevisiae_GO_matrix_wide
GO_table <- GO_table[-IndexRemoved,]
# Check if the genes were properly removed 
diff_GO.Count2 <- setdiff(GO_table$GeneID, rownames(agg))
dim(diff_GO.Count2)

# Save the new GO table
saveRDS(GO_table, "GO_table.rds")

```

### 10. Blind GO Terms (90-10 split)

30 per cent of GeneIDs will have zero Gene Ontology Annotations. This will serve as the training data. After the model has been optimised, these IDs will be unblided as the testing data.

```{r echo=TRUE}
#Fraction of the total number of genes to be blinded
test_size <- floor(0.10 * nrow(GO_table))

## set the seed to make your partition reproducible
set.seed(42)
blinded_ind <- sample(seq_len(nrow(GO_table)), size = test_size)
GO_train <- GO_table
GO_train[blinded_ind, 2:ncol(GO_train)] = 0

saveRDS(GO_train, "GO_train.rds")

```

### 11. Data preparation functions

```{r eval=FALSE, include=FALSE}

# Determine min-max cut value through trial and error to get 20 to 2000 total clusters

cuttree_val_ward <- cl_lengthCut(hr=hr, min=4, max=2000, interval=2)
saveRDS(cuttree_val_ward, "cuttree_val_ward.rds")

```


### Session Information 

```{r Session Info, echo=FALSE}

sessionInfo()

```
