---
title: "Yeast Gene Imputation (blinded data)"
author: "M.Soria"
date: "19/10/2020"
output: html_document
---
</br>

***

##### The following code imputes function on genes from Sacharromyces cerevisiae. Some of the genes were randomly chosen and "blinded" for testing recall.
***

</br>

### Prerequisite libraries
```{r Libraries, echo=TRUE}
suppressPackageStartupMessages({
    library(R.utils)
    library(dplyr)
    library(tidyr)
    library(data.table)
    library(RColorBrewer)
    library(gplots)
    library(Polychrome)
    library(tidyverse)
    library(caret)
    library(ggplot2)
    library(reshape2)
    library(tools)
    library (igraph)
})

source("functions.R")

```
</br>

### Load Pre-requisite RDS files 
```{r echo=TRUE}
file_PASS <- "tmm_agg_s.rds"
file_GO <-"GO_table.rds"
file_cuttree_values <- "cuttree_val_ward.rds"

tmm_agg_s <- readRDS(file_PASS)
GO_table <- readRDS(file_GO)
cuttree_values <- readRDS(file_cuttree_values)
```

### 1. Blinding known GO terms

```{r}
#Fraction of the total number of genes to be blinded
nfolds <- 10

set.seed(42)
folds <- sample(1:nfolds, nrow(GO_table), replace = TRUE)
#fix rownames
dict_folds <- data.frame(GO_table$GeneID, folds)

saveRDS(dict_folds, "dict_folds.rds")


# Make a data matrix with the blinded genes = 0
which_fold <- 3

ind_blinded <- which(dict_folds$folds == which_fold)
GO_blinded <- GO_table

GO_blinded[ind_blinded, 2:ncol(GO_blinded)] = 0

```

### 2. Coarse Sweep for Spearman + complete linkage (SandC)
The following code will perform a K-fold validation for the imputation process which uses Pearson method to measure the distance matrix and complete as linkage method during the clustering process. 

```{r eval=FALSE, include=FALSE}
# Coarse sweep of the following parameters: cluster total(20, 50, 200, 500) and threshold (0.1, 0.4, 0.8)

cl_list <- c(20, 50, 100, 300, 500)
thresh_list <- c(0.1, 0.4, 0.8)

CS_ward <- kfold(cl_list, thresh_list, cuttree_values, tmm_agg_s, GO_table, GO_blinded)
    saveRDS(CS_s, "CS_ward.rds")

# Testing bigger clusters

GO_blinded <-readRDS("GO_blinded.rds")
thresh_list <- c(0.1, 0.4, 0.8)

cl_list700 <- c(700)
CS_700_ward <- kfold(cl_list700, thresh_list, cuttree_values, tmm_agg_s, GO_table, GO_blinded)
saveRDS(CS_700_ward, "PredictionScores/CS_700_ward.rds")

```

### 3. Summary of performace measures for Pearson + complete linkage

```{r}

dat_list <- lapply(CS_ward, function(x) x[names(x)!="Index_folds"])

# List the all the mean predicton scores for each parameter 
summary_all <- summary_Csweep(dat_list)
saveRDS(summary_all, "summary_all.rds")

```

### 4. Visualizations

```{r}

# cut tree value vs Total Clusters – line diagram
cuttree_totals_df <- function(cuttree_values){
  df <- as.data.frame(as.numeric(names(cuttree_values)))
  m <- list()
  for (i in 1:length(cuttree_values)){
    x <- cuttree_values[[i]][["Cut_value"]]
    m[[i]] <- x
  }
  df$CuttreeVal <- as.numeric(m)
  colnames(df)[1] <- "ClTot"
  png("Data/cuttreeVSclTotal.png",
    width = 5*300,        # 5 x 300 pixels
    height = 5*300,
    res = 300,            # 300 pixels per inch
    pointsize = 8)        # smaller font size
  
  ggplot(df, aes(x=CuttreeVal, y=ClTot)) +
    geom_line() + ylab("Total Number of Clusters") +
    xlab("Cuttree Values") +
    ggtitle("Cuttree Value vs Total Number of Clusters") +
    theme(plot.title = element_text(hjust = 0.5))
  
  dev.off()
}

cuttree_totals_df(cuttree_values)

 


# Total genes per cluster  – violin plot
c_violin <- function(cuttree_values){
  
  clTotals <- do.call(rbind.data.frame, sapply(cuttree_values, "[[", 3))
  clTotals$ClusterSize <- as.factor(as.numeric(str_extract_all(rownames(clTotals), "^[:digit:]*(?=.)")))
  clTotals$ClusterID <- as.numeric(str_extract_all(rownames(clTotals), "(?=.)\\d+$"))
  colnames(clTotals)[1] <- "TotalGenes"
  rownames(clTotals) <- c()
  
  # Filter according to cluster size
  cluster20_100 <- clTotals[clTotals$ClusterSize %in% c(20,50,100),]
  cluster200_500 <- clTotals[clTotals$ClusterSize %in% c(200, 300, 500),]
  cluster800_1500 <- clTotals[clTotals$ClusterSize %in% c(800, 1000, 1500),]
  
  pdf("Data/Violinplots.pdf")
  
  ggplot(cluster20_100, aes(x=ClusterSize, y=TotalGenes, fill=ClusterSize)) +
    geom_violin(trim=FALSE) + ylab("Total Number of Genes") +
    xlab("Cluster Size") +
    ggtitle("Distibution of genes per cluster size (20 to 100)") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_boxplot(width=0.1, fill="white") + theme_minimal() +
    scale_fill_brewer(palette="Blues")
    
  ggplot(cluster200_500, aes(x=ClusterSize, y=TotalGenes, fill=ClusterSize)) +
    geom_violin(trim=FALSE) + ylab("Total Number of Genes") +
    xlab("Cluster Size") +
    ggtitle("Distibution of genes per cluster size (200 to 500)") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_boxplot(width=0.1, fill="white") + theme_minimal() +
    scale_fill_brewer(palette="Dark2")
  
  ggplot(cluster800_1500, aes(x=ClusterSize, y=TotalGenes, fill=ClusterSize)) +
    geom_violin(trim=FALSE) + ylab("Total Number of Genes") +
    xlab("Cluster Size") +
    ggtitle("Distibution of genes per cluster size (800 to 1500)") +
    theme(plot.title = element_text(hjust = 0.5)) +
     geom_boxplot(width=0.1, fill="white") + theme_minimal() + 
    scale_fill_brewer(palette="RdBu")
  
  dev.off()
}

c_violin(cuttree_values) 


```

### Imputation review

```{r}

# CL Tot = 50, thresh = 0.8

impute <- function (cl_GOall, corr_clAll, clust_total, thresh){
  wGO_list <- list()
  
  for (i in 1:clust_total){
    
    GO_list <- rownames(cl_GOall[[i]])
    
    #If all the genes in the cluster have no GOs and the matrix is empty
    if (is_empty(GO_list) == TRUE){
      wGO_list[[paste0("Cluster", i)]][[paste0("Comment")]] <- "No Gene Ontologies found"
      wGO_list[[paste0("Cluster", i)]][[paste0("Input")]] <- 0
      wGO_list[[paste0("Cluster", i)]][[paste0("Output")]] <- 0
      wGO_list[[paste0("Cluster", i)]][[paste0("Diff")]] <- 0
      
      next
    }
    
    GO_list <- GO_list[order(GO_list)]
    
    corr_cl <- corr_clAll[[i]]
    corr_cl <- corr_cl[order(rownames(corr_cl)),]
    corr_cl <- corr_cl[,order(colnames(corr_cl))]
    gene_list <- rownames(corr_cl)
    
    cl_go <- cl_GOall[[i]]
    
    #If all the genes in the cluster have no GOs and the matrix is empty
    if (is_empty(cl_go) == TRUE){
      wGO_list[[paste0("Cluster", i)]][[paste0("Comment")]] <- "No Gene Ontologies found"
      wGO_list[[paste0("Cluster", i)]][[paste0("Input")]] <- 0
      wGO_list[[paste0("Cluster", i)]][[paste0("Output")]] <- 0
      wGO_list[[paste0("Cluster", i)]][[paste0("Diff")]] <- 0
      
      next
    }
    
    # should add gene that are not included in the GO data otherwise merge 
    # will yield weird results
    diff_go_genes <- setdiff(gene_list,rownames(cl_go))
    add <- data.frame(matrix(0,nrow=length(diff_go_genes),ncol=ncol(cl_go)))
    rownames(add) <- diff_go_genes
    colnames(add) <- colnames(cl_go)
    cl_go <- rbind(cl_go,add)
    cl_go <- cl_go[order(rownames(cl_go)),]
    cl_go <- cl_go[,order(colnames(cl_go))]
    
    wGO_cl <- wcorr_cluster(gene_list=gene_list, corr_cl=corr_cl, cl_GO=cl_go)
    wGO_df <- as.data.frame(do.call(rbind, wGO_cl))
    wGO_thresh <- (as.matrix(wGO_df) > thresh)*1 
    wGO_thresh <- wGO_thresh[order(rownames(wGO_thresh)),]
    wGO_thresh <- wGO_thresh[,order(colnames(wGO_thresh))]
    
    cl_subtract <- wGO_thresh - cl_go
    # need to flip some of the values where NAs were replaced with zeros.
    cl_subtract[cl_subtract<0] <- 1
    
    wGO_list[[paste0("Cluster", i)]][[paste0("Input")]] <- cl_go
    wGO_list[[paste0("Cluster", i)]][[paste0("Output")]] <- wGO_thresh
    wGO_list[[paste0("Cluster", i)]][[paste0("Diff")]] <- cl_subtract
    
  }
  return(wGO_list)
}

cross_val <- function(n, GO_annot, clusters,
                      GO_list_perCl, corr_clAll,
                      wGO_allClusters, clust_total, 
                      thresh){
  stats <- list()
  
  # partition the data
  nfolds <- n
  set.seed(42)
  folds <- sample(1:nfolds, nrow(GO_annot), replace = TRUE)
  
  dict_folds <- data.frame(GO_annot$GeneID, folds)
  
  for (i in 1:n){
    
    which_fold <- i
    
    ind_blinded <- which(dict_folds$folds == which_fold)
    GO_blinded <- GO_annot
    GO_blinded[ind_blinded, 2:ncol(GO_blinded)] = 0
    
    # Get the GO terms by cluster using the blinded data 
    # and the GO list per cluster
    GO_blindedCl <- GO_per_cl_blinded(GO_blinded, clusters, 
                                      GO_list_perCl, clust_total)
    
    # Impute function
    wGO_blinded <- impute(GO_blindedCl, corr_clAll,
                          clust_total, thresh)
    
    # stats
    stats_perCl <- stats_cl(wGO_allClusters, wGO_blinded, clust_total)
    stats_final <- stats_all(stats_perCl)
    
    stats[[paste0("Fold", i)]] <- stats_final
  }
  stats[["Index_folds"]] <- dict_folds
  return(stats)
  
}

kfold <- function(cl_list, thresh_list, cuttree_values, logcounts=tmm_agg_s, GO_table, GO_blinded){
  
  scores <- list()
  
  for (i in cl_list){
    print(i)
    for (j in thresh_list){
      print(j)
      cl_tot <- as.character(i)
      cl <- cuttree_values[[cl_tot]][["GeneID_assignments"]]
      
      corr_allCl <- corr_per_clust(logcounts, cl, i)
      GOterms_perCl <- GO_per_cl(GO_table, cl, i)
      GO_list_perCl <- GO_list_perClBlind(GOterms_perCl, i)
      
      # stats for thresholds
      wGO_allCl <- impute(cl_GOall = GOterms_perCl,
                          corr_clAll = corr_allCl,
                          clust_total = i, thresh = j)
      
      sc <- cross_val(n=10, GO_annot=GO_table, clusters=cl,
                      GO_list_perCl=GO_list_perCl,
                      corr_clAll=corr_allCl,
                      wGO_allClusters=wGO_allCl,
                      clust_total=i, thresh=j)
      
      scores[[paste0(i, "_", j)]] <- sc
    }
  }
  return(scores)
}

```


### Session Information 

```{r Session Info, echo=FALSE}
sessionInfo()
```
